\startcomponent 00_condorcets_father

\scene

\open The muffled chanting of La Prise de La Bastille can be heard from some
other room. After a time, the song comes to a halt as \condorcet is revealed
holding his infant daughter.

\stage Pause.

\startcondorcet
\(to his daughter)
I never knew my father.

\(beat)
I've been told he was a noble chevalier, a soldier of principle and honor, a
warrior from a lineage of warriors; bound by the oath he swore to the motto of
our family.

A motto consisting of just a single word: "caritas." Charity.

\(beat)
But I never knew him.

By the time I was your age, my dearest \eliza, my father had perished.

Leaving me no fortune but pride, and no teaching but that single word.

\(beat)
I was raised alone by my mother, who was crippled by the weight of her
responsibility as a widow and perpetually haunted by her awareness of her own
ignorance.

She was as tender as any son could ask, but her anxiety and superstition took
the best of her.

% TODO: reword.
In fact, until I reached the age of nine, my mother had me dress as a girl each
and every day - believing that having me wear skirts was an act of devotion to
the virgin Mary, and that the virgin's blessing would protect me from all the
evils of an uncertain universe.

\(beat)
And yet... despite how humiliating her possessiveness had been, I learned early
on to forgive her.

Her zealous paranoia had always been rooted in her vast capacity for love.

It was merely the consequence of tragic circumstances that led her to distort
that love the way she had.

I knew that she was not to be blamed for such circumstances, and thus... I
learned how to forgive.

\(beat)
The most formative experience of my life was when I first realized just how
afraid she was of all that which she knew that she did not know.

But that, dear \eliza, shall be the past.

It is my greatest pleasure to have brought you into this world at the perfect
moment.

You will belong to the first generation of women to receive the same education
as men and participate in the sciences as their equal.

\(beat)
Never again shall any woman suffer as she had.
\stop

\instruction The chanting returns.

\stopcomponent
