\startcomponent assembly

\scene

\open The Salle du Man\`ege - the makeshift auditorium that's clearly
unsuitable for a formal government.

\stage Enter \condorcet, followed by \leibniz.

\startleibniz
Ugh, this is truly where the first legislative assembly of France shall meet?

Unacceptable!

The geometry is all wrong! The proportions of this hall are far too narrow and
the acoustics cannot be tolerated.
\stop

\stage Enter \voltaire.

\startvoltaire
What? Not a fan of needing to turn your neck ninety-degrees to see who's
speaking?

Or needing to cup your hands behind your ears to hear them?
\stop

\stage Enter the \deputies, including \brissot.

\stage Working-class citizens, including \robespierre, begin filling the
galleries.

\direction \robespierre is famous among the women of Paris; a group of
lower-class women accompany him in the galleries.

\stage One \deputy seems to be lost.

\startvoltaire
\(approaching the \deputy)
Aw. Look at this poor boy. The man seems completely out of his element here -
his eyes dart around the room looking for any semblance of connection to this
world.

\(to \condorcet)
Where do you suppose he should sit? To the left? Or to the right? Oh, god, how
perilous this choice must seem to him. How suffocating its consequences!
\stop

\startleibniz
Leave the man to his peace, \voltaire. This assembly is full of inexperienced
politicians and we ourselves are no different.
\stop

\startvoltaire
Nothing but greenhorns as far as a twist of the neck can tolerate.

\(gestures to \robespierre)
Ah! And look! There he is: the man responsible for this dilemma comes to admire
his handywork.
\stop

\stage \robespierre takes out his notebook.

\startleibniz
Indeed, it is quite curious why someone as shrewd as \robespierre would
use his influence to prevent the members of the National Assembly from
participating in France's new constitutional monarchy.
\stop

\stage \voltaire chuckles - shaking his head.

\startvoltaire
Oh, whatever shall we do with you, \leibniz?
\stop

\stage Exit \voltaire and \leibniz.

\stopcomponent
