\startcomponent aux_armes_2

\scene

\stage \condorcet is startled. \sophie giggles.

\startcondorcet
Huh, what's this about?
\stop

\startbrissot
We were only discussing how rude it is for philosophers to ignore their
company.
\stop

\stage \sophie punches \brissot in the arm. \brissot laughs.

\startpaine
He is certainly braver than I am in ignoring conversations that bore him.

\(chuckles)
I should do it more often.
\stop

\startsophie
Monsieur \paine would like to know when you plan to give your speech to the
\crowd[jacobins].
\stop

\startcondorcet
Ah! It shall be next Friday.
\stop

\startbrissot
I still insist that the climate is not yet right for such a bold display.
\stop

\startcondorcet
The timing is perfect - the king has shattered his own crown! Why wouldn't we
pursue this opportunity?
\stop

\startbrissot
I'm not against the idea of you exercising your free speech, but what I'm
suggesting is that it would be best to wait until after the coming elections.

At the moment, tensions are high.

\barnave and \lafayette know that their position is weak and they feel
themselves backed into a corner.

Nonetheless, together they still maintain a strong majority in the assembly,
meaning you'd be putting your campaign at risk for no strategic benefit.
\stop

\startpaine
What kind of risk?
\stop

\startbrissot
Well, assuming you don't incite a riot, you'll have the royalist propaganda to
deal with.
\stop

\startsophie
Oh, heavens...
\stop

\startbrissot
Mhm, I'm afraid so.

The royalist presses are loud, crude, vicious, and worst of all - shameless.

My guess is that they're waiting for someone to speak up on the republic so
they can make an example of them in their broadsheets.
\stop

\startcondorcet
Ha! If it's bullies that have you afraid, I shall tell you that I am no
stranger.

As a boy, my mother forbade me from wearing my breeches until I had reached the
age of nine.

I assure you that the cruelty of child-like minds are not a burden to me.
\stop

\stage \paine and \brissot stifle their laughter. \sophie punches each in the
arm.

\startbrissot
Children or no, they are well practiced in the art of baseless insults.
\stop

\startpaine
\condorcet is right. We shouldn't cower at a time like this.

The king has insulted the people by abandoning his duty - to my mind, that
renders him abdicated.

We should consider starting a journal for the cause of a republicanism.
\stop

\startsplit

\startcondorcet
Yes!
\stop

\split

\startsophie
Yes! Wonderful idea!
\stop

\stopsplit

\startbrissot
Hmm...
\stop

\startpaine
\brissot?
\stop

\startbrissot
Right now, our focus should be on deposing the king.
\stop

\startpaine
Then we share a common goal.
\stop

\startbrissot
Not exactly. There's a difference between replacing a king and replacing a
monarchy.

I'm afraid that a republic will require something much more dramatic than a
journal.
\stop

\startpaine
Oh, is that so? Well, you know what has me afraid? I'm afraid you're hiding
something from us.
\stop

\startbrissot
Oh?
\stop

\startcondorcet
What's this? Are you planning something, \brissot?
\stop

\startbrissot
\(chuckles)
In due time, gentlemen. In due time.
\stop

\startcondorcet
You are! Why be secretive? Why not trust your allies?
\stop

\startpaine
He must suspect us of being politicians.
\stop

\stage \brissot laughs.

\startbrissot
I would never!
\stop

\stage \paine laughs.

\startsophie
Now that's been mentioned, there has been something that has kept me suspicious
of you.
\stop

\startbrissot
Oh? What would that be, Madame?
\stop

\startsophie
In your little speech earlier, you spoke as if you had the whole show
rehearsed.
\stop

\startbrissot
I take that as a complement! I'm proud of my talent for speaking from the
heart.
\stop

\startsophie
Yet, that's what has me confused. Your speech would have fallen flat had it not
been for the drunkard playing along.

\(beat)
How did you know he would provide your punchline?
\stop

\stage \brissot is impressed.

\startbrissot
You're insinuating I wouldn't take a gamble like that on a whim.

\(beat)
Alright, such a sharp eye deserves the respect of honesty.

The truth is, I had paid the man prior to our arrival.
\stop

\startcondorcet
What? You paid him? Why go through such an effort for a simple laugh?
\stop

\startpaine
Gossip.
\stop

\startbrissot
Ah, an astute learner! He may become a proper Frenchman yet.
\stop

\startcondorcet
Gossip?
\stop

\startsophie
What kind of gossip?
\stop

\stage \brissot clears his throat, then stands on his chair. The
\crowd[patrons] turn to look at him.

\startbrissot
TO ARMS! TO ARMS!
\stop

\startcrowd[cue={Patrons}]
TO ARMS! TO ARMS!
\stop

\stage The \crowd[patrons] laugh.

\startbrissot
All the world's a stage, my friends.
\stop

\direction An infant begins to cry from the other room.

\stopcomponent
