\startcomponent brissotins

\scene

\open Spring 1792. The home of \vergniaud. The \brissotin[party] have gathered
for dinner.

\startvergniaud
So, what will it be tonight? What are you fiends in the appetite for?
\stop

\stage \brissot is lost in contemplation.

\startsplit

\startguadet
Give us \marat!
\stop

\split
\leavevmode

\startisnard
No! No! \condorcet!
\stop

\stopsplit

\startvergniaud
No. I can't stand doing \marat. Even in jest he gives me a migraine.
\stop

\startsplit

\startguadet
Oh, don't be such a stooge!
\stop

\split

\startgensonne
\robespierre!
\stop

\stopsplit

\startvergniaud
\robespierre? Ah, a classic!
\stop

\stage \vergniaud switches his posture to that of a predator as he cartoonishly
stalks the other \brissotins like prey.

\stage Pause. Silence. Minus a few chuckles.

\startvergniaud
To arms? What's this business of arms? Whose arms? How many arms?

What do these arms do? And why?
\stop

\stage Laughter. \vergniaud remains deadly serious.

\startvergniaud
Consider, citizens, the arms of our enemies.

They must have two for every soldier, but how do we know they don't possess
more?

\beat

The Romans had arms, yet consider their fate and ask yourselves whether our
arms should truly fare any better?

With my anxious glare, the arms I see are only suited for gorging on bread and
composing pornography.

\beat

I assure you, citizens, that if I were the master of all the arms in France, I
would have long ago formed them into a giant arm-monster capable of devouring
any who stand opposed to virtue.
\stop

\stage Laughter. \vergniaud joins in this time.

\startguadet
Oh, come now, \brissot.

Don't tell me you're too bitter about your spat with \robespierre to laugh at
his expense!
\stop

\startbrissot
No, gentlemen. I'm only thinking.
\stop

\startvergniaud
My friend, in the time I've known you, not once have I seen you take a break
from thinking.
\stop

\startbrissot
\beat{chuckles}

I suppose that's true, \vergniaud. Must be \condorcet's influence on me!
\stop

\stage Laughter.

\startbrissot
But, no. I contemplate grim observations that I've made of the assembly.
\stop

\startguadet
You've only now begun to notice? You're a bit slow if that's the case, I'm
afraid.
\stop

\startgensonne
He's just bitter about losing the election for president again!
\stop

\startbrissot
I suppose you could say that, though my bitterness is not selfish, I assure
you.

All that matters is your own victory, \gensonne.

I am honored to be merely a supporter.
\stop

\stage \brissot stands up to begin pacing the dining room.

\startbrissot
Nonetheless, I find that losing the election for president by such a narrow
margin - by the {\emph same} margin - for each of the nine ballots that bore my
name...

I believe this is a symptom of a larger issue.
\stop

\startguadet
Well, get on. What's this about?
\stop

\startbrissot
I fear that the assembly has fully embraced factionalism.

It's clear to me that royalists have successfully formed a conservative party
to oppose our agenda and collude with the crown.
\stop

\stage The \brissotins exchange stern glances.

\startbrissot
I say, gentlemen, that if we are up against collusion, then we have no choice
but to respond in kind.
\stop

\startisnard
I won't debase myself by submitting to factionalism!
\stop

\startbrissot
I commend your principles, my friend.

I, too, feel disgusted with the mere thought of it.

But forcing the court to declare war will be impossible while the conservatives
gather influence below the noses of the sane.
\stop

\startvergniaud
If I may be so bold, I believe your grievances are overblown, \brissot.
\stop

\startbrissot
How so?
\stop

\startvergniaud
We have the support of the public. The people of France are unified in their
desire for war. No collusion can penetrate this fact.
\stop

\startsplit

\startbrissot
Perhaps you have a point...
\stop

\split

\startgensonne
Speak for yourself, \vergniaud.
\stop

\stopsplit

\startguadet
But what of \robespierre?
\stop

\startvergniaud
What of him?
\stop

\startbrissot
No. \robespierre is my own burden to bear.

The feud between he and I is personal, and I'm confident that the public
understands this.
\stop

\startisnard
How can you be sure? \robespierre has effectively declared himself the king of
the Jacobins.
\stop

\startguadet
Doubtless \robespierre will stop at nothing to declare himself the king of the
French.
\stop

\stage The \brissotins grumble at the thought.

\startbrissot
Sadly, I agree. \robespierre thinks he alone speaks with the voice of the
people.

What ego. What insufferable ego.
\stop

\startisnard
Hear, hear!
\stop

\startbrissot
But he is not a legitimate threat to us.

Gentlemen... we are the ones in power, not him. Let us not forget this.
\stop

\startvergniaud
Then what say you? What's our next step?
\stop

\startgensonne
Next steps? Why do we rush while our enemies hesitate? I agree with \condorcet
on the matter of war.

Economic solutions are what France needs.

Our true enemy is starvation, and if we are wrong about swift success, if our
borders falter even slightly, then our enemies will be sure to destroy our
grain reserves.
\stop

\startvergniaud
Spoken like a man with true love for his country!
\stop

\startbrissot
Hear, hear! This perspective speaks well of you, \gensonne.

In fact, under most circumstances, I would agree.
\stop

\startgensonne
Is France not in such circumstances?
\stop

\stage Pause.

\startbrissot
It is with great pain that I must say no.

\beat

\condorcet fails to see that ideals belong only to enlightened nations...

Yet, France, I fear to say... is not such a nation.

\beat

Prejudice, not reason, has become the dominant force of our hearts...

Prejudice that has been fostered by our idleness.

\beat

Gentlemen, consider how much we have suffered the voice of fools and fiends.

Men who dare call themselves representatives of humanity, who stand on the
tribune to argue for the continued enslavement of our brothers of color.

Consider the hate that has overrun our streets and the popular vengeance that
ignorant men wield against the innocent.

Consider the denunciations that have inflamed all our journals.

CONSIDER... how much {\emph damage} that slander and libel has done to even our
most treasured of human bonds.

\beat

No.

Bless \condorcet for his na\"ivety.

Bless him for his love of reason, and bless the battle he wages on the front of
education.

But gentlemen, we need more than reason. We... need action.

Action is what will win the hearts of the ignorant and destroy the disease of
prejudice.

And, gentlemen, there is no greater action than that of war.

\beat

War alone has power that is sufficient to compel every man who loves their
country to at once forgive their neighbors.

To see in them not as schemers, or intriguers, or enemies of liberty...

\beat

But as comrades... and as brothers...
\stop

\startvergniaud
TO ARMS, THEN!
\stop

\startbrissotins
TO ARMS! TO ARMS!
\stop

\stage The \brissotins applaud \brissot.

\stopcomponent
