\startcomponent cordeliers_3

\scene

\open Summer 1792. Home office of \danton.

\stage Enter \robespierre.

\startdanton
Ah! There's the haunt of Paris!
\stop

\startrobespierre
Good evening, \danton.
\stop

\stage \robespierre takes a seat as \danton pours him a drink. The glass is
never touched.

\startdanton
So. What's the rush to get a meeting with me?
\stop

\startrobespierre
I wished to discuss with (you) --
\stop

\startdanton
-- You're lucky, by the way.

\(beat)
If my buddy Camille didn't insist that you were truly, deep down, a cuddly
little kitten - I would not have made the time for this meeting.
\stop

\startrobespierre
Camille is a childhood friend of (mine) --
\stop

\startdanton
-- You know.

\(deep breath)
This war, brother; this war. It's really eating me up.

\(tsk tsk tsk)
I think it has really begun to take its toll on France, wouldn't you agree?
\stop

\stage Pause.

\startdanton
Something must be wrong. \louis is up to something. He must be~--
\stop

\startrobespierre
-- I agree. It's obvious that \louis would wish to sabotage the war with
Austria.

Allowing the enemy to march into Paris to take our heads allows him to restore
the monarchy.

\(beat)
But I'm afraid that it is not \louis who France must fear.
\stop

\stage Pause. Silence.

\startrobespierre
I believe that France's true enemies are the moderates in the assembly who have
clumsily thrown us into this war.

Those fools who wear the mask of patriotism to hide their idiocy.
\stop

\stage \danton laughs.

\startdanton
You mean \brissot? Aw! Not monsieur patriot!
\stop

\stage \danton chuckles.

\stage Pause. Silence.

\startrobespierre
I hate him. I hate him more than you can know.
\stop

\startdanton
What do you want me to do about it.
\stop

\startrobespierre
I'd like to know how you feel about these deputies.
\stop

\startdanton
I feel how I will about them. But right now, you're in my office.

And while France is at war, feelings like these tend to come at a lofty cost.

Let's start with what you make of the puppy that follows them around.
\stop

\startrobespierre
You mean \condorcet?

I admit. I have been impressed with his ideas.

I'm particularly fond of his stance on allowing women to divorce their
husbands.
\stop

\stage \danton[thedaring] squares up to \robespierre[theincorruptible].

\stage Pause.

\startrobespierre
Many of my supporters are women.
\stop

\startdanton
I've noticed. But what's the point of having women support you? Women don't
vote.
\stop

\startrobespierre
Their husbands do.
\stop

\stage \danton takes a second, then clicks!

\startdanton
Oh, FUCK!

HA! That's brilliant!
\stop

\stage \danton heartily laughs.

\stage \danton collects himself - he is delighted.

\startdanton
I suppose you have the philosopher to thank. HA!

I doubt you'd be brave enough to announce on the tribune that you're a cuckold!
\stop

\startrobespierre
I disagree. One must only say it correctly - and under the correct
circumstances.
\stop

\stage Pause. A smug suspense.

\startdanton
You really are something else, aren't you?
\stop

\stage \robespierre stands to exit.

\startdanton
And you're finished?
\stop

\startrobespierre
I only requested a brief moment of your time. I wish to honor this.
\stop

\stage Exit \robespierre.

\stage \danton takes \robespierre's drink and begins casually pacing the room.

\stage \danton becomes suspicious.

\startdanton
What the hell just happened?
\stop

\stage \danton puzzles on his thoughts.

\stage He shakes his head, downs his drink, and takes a deep breath.

\startdanton
No... no... damn it. He's right.
\stop

\stopcomponent
