\startcomponent cordeliers

\def\startone{\startsansculotte[number=1]}
\def\starttwo{\startsansculotte[number=2]}
\def\startthree{\startsansculotte[number=3]}
\def\startfour{\startsansculotte[number=4]}

\scene

\open The Convent of the Cordeliers. A crowd of \sansculottes are stirring in
anticipation.

\startone
I hear that she's a vampire!
\stop

\startsplit

\starttwo
A vampire?
\stop

\split

\startthree
What's a vampire?
\stop

\stopsplit

\startone
It's a kind of demon that feeds by killing folk and drinking their blood.
\stop

\starttwo
I heard she was one of those incest types? From when she was an Austrian?
\stop

\startthree
She's still an Austrian, ya nit!
\stop

\starttwo
But back when she was {\emph in} Austria. I heard that she would plow her
brothers.
\stop

\startone
Plow her brothers?
\stop

\startthree
Yeah, emperor Leopold and, uh, Joseph and the like - back when they was kids.
\stop

\startone
Probably why she and her mule tried to flee back home... for nostalgia.
\stop

\startfour
Ey! They're here! They're here!
\stop

\stage Enter \marat and \danton.

\stage The \sansculottes are riotous.

\startsplit

\startsansculottes
\marat! \marat! \marat!
\stop

\split
\leavevmode

\startone
The friend of the people!
\stop

\stopsplit

\stage \marat stands on a makeshift podium.

\startmarat
THE NATIONAL ASSEMBLY IS CORRUPT! THEY ARE A CABAL OF FIENDS!
\stop

\stage The \sansculottes enter a frenzy!

\startmarat
The cowards at the assembly have just decided to offer the king and his whore
amnesty for their attempt to wage a war against the revolution!

Led by his Austrian spy, \louis[capet] - this man who has made himself a
common bastard - wants nothing less than to lead a foreign army into Paris so
that they may rule over twenty-five million corpses!

And our corrupt deputies at the National Assembly have just pardoned them!

They wipe their asses with the rights of man!

\(beat)
We shall see France hand over the power of the army to this man and his lackeys
who would trample upon the innocent, the poor, the starving - they would steal
the bread from your stomachs out of spite!

They spit in your face out of disgust for men who demand NOTHING BUT THEIR
DIGNITY!

\(squares up for a fight)
This day last year, you would remember that I told you five hundred heads would
have sufficed; today, fifty thousand are necessary!

And if we don't act, by the end of the year, France shall need FIVE HUNDRED
THOUSAND HEADS!

Paris would be flooded with blood and not be any freer for it!
\stop

\stage \danton's bellowing laugh can hardly be heard over the commotion.

\startsplit

\startsansculottes
TO THE ASSEMBLY!
\stop

\split

\startsansculottes
WE START TODAY!
\stop

\stopsplit

\startdanton
BROTHERS! BROTHERS!
\stop

\stage The \sansculottes simmer down.

\starttwo
Quiet! \danton speaks!
\stop

\stage Pause. A smug suspense.

\startdanton
Let's not lose our own heads before we begin taking theirs!
\stop

\stage Laughter.

\startdanton
\marat's feverish spirit speaks to all our hearts, but we must act with the
same care as with all our labors!

Think of our women! We would frighten them to return coated in the blood of
reckless abandon.

Instead, we must save our barbarity for the sheets, but only after we impress
them with our civility!

\(beat)
Gentlemen, I happen to know that all of Paris feels much the same way we do
about the king's treachery.

We still yet have hope for our patriotism to shine in the public fervor.

I say we petition the \jacobins for their support in overturning this vote and
having this king tossed!
\stop

\stage Pause. Suspense.

\startdanton
Failing that, then we'll start shaving the bastards!
\stop

\stage \danton laughs. The \sansculottes cheer.

\stopcomponent
