\startcomponent cordeliers_2

\scene

\open Cellar of the Cordeliers club. \marat sits in a medicinal bath while he
writes.

\stage Enter a \woman.

\startwoman
\marat, you have a visitor.
\stop

\startmarat
Who?
\stop

\startwoman
A man by the name of \robespierre.
\stop

\stage Pause.

\stage \marat nods to the \woman. She exits.

\stage Enter \robespierre.

\stage Pause.

\startmarat
What do you want?
\stop

\startrobespierre
I've come to interview you.
\stop

\startmarat
For what paper?
\stop

\stage \robespierre procures a small notebook.

\startrobespierre
For my notes.
\stop

\startmarat
Why?
\stop

\startrobespierre
Because of every man I have ever known, you are the first that I do not
understand.

I cannot allow this.

\(beat)
When that fatal petition was brought to the \jacobins, it was the terror you
inspired in my rivals that led to the defeat of my efforts to prevent the
deaths of two hundred innocent civilians.
\stop

\startmarat
Two thousand innocents!
\stop

\startrobespierre
Far too many to bear.
\stop

\startmarat
You dare come here to accuse me~--?
\stop

\startrobespierre
-- No. Your actions were reasonable.

They adequately expressed the general will.

\(beat)
Nonetheless, I ask for your sympathy regarding my position.

You see, I cannot allow someone I do not understand to influence the decisions
necessary for public safety.

\(beat)
Are you willing to accept the interview?
\stop

\stage \marat sneers.

\startrobespierre[continued]
I understand that you are ill?
\stop

\startmarat
The bath is medicinal. It soothes my skin irritation.

\(beat)
You're lucky it's not contagious.

Next time you insist on speaking to someone you suspect of being ill, bring a
cloth to cover your face with.
\stop

\startrobespierre
You practice medicine?
\stop

\startmarat
I used to, I switched to studying physics when I realized medicine didn't let
me exercise my true curiosity for the universe.
\stop

\stage Pause.

\startmarat[continued]
Don't bother looking for my work anywhere near the Academy.

They refused to publish anything from the thousands of experiments I have done.

Not a one.

\(spits)
Those elitists called me a quack for suggesting that Newton could have been
wrong about something.
\stop

\startrobespierre
Did you discover a contradiction with one of Newton's experiments?
\stop

\startmarat
No. I just don't believe that it's possible to model the universe with numbers.
\stop

\startrobespierre
I have noticed that, in your speeches, you always seem to demand a specific
number of heads, yet I notice that you never demand a consistent quantity.
\stop

\startmarat
The more heads I demand, the less serious I'm being.
\stop

\stage Pause. Silence.

\startrobespierre
I also notice that you never provide any names for who these heads should
belong to.
\stop

\stage Pause. Silence.

\startrobespierre[continued]
You do not know who your enemies are.

\(beat)
You certainly know that they exist, but not what they look like or where they
will strike from.

You are hunted like a rabbit with its eyes gouged and ears punctured.

You can smell your predators, you know they are near, yet this does not terrify
you?
\stop

\startmarat
I don't fear dying for the revolution.
\stop

\startrobespierre
Why not?
\stop

\startmarat
Why not? You think I should take the shit they do in silence?

These fiends that would speculate on the price of bread to profit from
starvation?

So they can collect shiny coins - or, god help me, paper - and enjoy the
pleasure of counting their collection?

\(beat)
You think just because I don't know who they are that they shouldn't be put to
death for their crimes?

You think that just because I haven't personally starved to death that my fate
should be any better than those who have?
\stop

\startrobespierre
I don't understand your tactics.

You threaten powerful people without considering the outcomes and without
considering the interests of --
\stop

\startmarat
-- No! I refuse to let the interests of those whoresons influence me - even if
it ends up getting me put down like a mad dog.
\stop

\startrobespierre
But why call for such extremes when they only doom your cause?
\stop

\startmarat
You obviously aren't going to understand. Get out.
\stop

\stage Pause.

\startrobespierre
I must understand.
\stop

\stage Pause.

\startmarat
Then understand that my reputation with the people does not rest upon my ideas.

It rests upon my boldness, upon the impetuous outbursts of my soul, upon my
cries of rage, of despair, and of fury.

I am their anger - their justified anger - and that is why they listen to me;
why they believe in me.

\(beat)
You think that I say these things because I'm just some fool who doesn't know
how to play your games?

You think that I have some agenda?

Some scheme that these calls for violence advance?

\(spits)
Understand that my cries of alarm that you take for empty words are the most
na\"ive and sincere expressions of the passions that devour my soul.
\stop

\stage \robespierre has discovered a new axiom~- causing him pause as he
considers its implications.

\stage \robespierre[theincorruptible] puts away his notebook and stands.

\stage Pause. Silence.

\startrobespierre
I believe you.
\stop

\stage \robespierre prepares to exit, then hesitates:

\startrobespierre[continued]
We should see about getting your listeners into the National Guard.

This is where true patriots belong, wouldn't you agree?
\stop

\stage Exit \robespierre.

\stopcomponent
