\startcomponent danton_negotiation

\scene

\open Summer 1792. Home office of \danton.

\stage Enter \robespierre.

\startdanton
Ah! There's the haunt of Paris!
\stop

\startrobespierre
Good evening, \danton.
\stop

\stage \robespierre takes a seat as \danton pours him a drink. \robespierre
never touches the glass.

\startdanton
So. What's the rush to get a meeting with me?
\stop

\startrobespierre
I wished to discuss with (you) --
\stop

\startdanton
-- You're lucky, by the way.

\beat

If my buddy Camille didn't insist that, deep down, you were truly a cuddly
little kitten, I would not have made the time for this meeting.
\stop

\startrobespierre
Camille is a childhood friend of (mine) --
\stop

\startdanton
-- You know.

\beat{deep breath}

% TODO: rewrite.
This war, brother; this war. Three months in and the enemy rampages through our
borders. Burning our grain - raping our women... Fucking. Disaster.
\stop

\stage Pause.

\startdanton
You know what I think? I think \louis is up to something. He must be~--
\stop

\startrobespierre
-- I agree. It's obvious that \louis would wish to sabotage the war to advance
his schemes.

Allowing the enemy to march into Paris to take our heads allows him to restore
the monarchy.
\stop

\startdanton
This shit's bad, brother. I admit, you were damn right about the whole thing
all along.
\stop

\startrobespierre
True. However, I believe that it is a mistake to assume that it is \louis who
we must fear.

I believe France's true~--
\stop

\stage \danton laughs.

\startdanton
-- You mean \brissot? Aw! Not mister patriot!
\stop

\stage \robespierre stands from his chair - offended.

\stage \danton instinctively grips the pistol in his coat.

\stage \robespierre turns and slowly begins pacing the office.

\startrobespierre
\brissot has invented a new form of evil that history has failed to prepare us
for.

He clumsily brandishes his mask of patriotism to obscure his incompetency and
bolster his insatiable ego.
\stop

\startdanton
\beat{chuckles}

Listen. I have no interest in your personal feud with mister patriot.
\stop

\startrobespierre
Personal? I assure you that I have no ambitions for personal satisfaction.

\beat{deep breath}

You see... \danton...

I have come to you today because I feared that you too have been led astray by
his chicanery.
\stop

\startdanton
\beat{squares up}

Oh, is that so?
\stop

\stage Pause.

\startrobespierre
I understand you have gained considerable influence over the National Guard.

Yet, from what I see, they nonetheless remain idle.
\stop

\startdanton
Idle? Fuck off! If you haven't noticed, France is defending itself against a
fucking siege.
\stop

\startrobespierre
Meanwhile, this force could instead be used for the offensive.
\stop

\startdanton
I have no intention of marching to Verdun and leaving the city defenseless.
\stop

\startrobespierre
You fail to understand my meaning. Consider instead the enemy already present
in Paris.
\stop

\startdanton
No. I will not send the Guard to slaughter the assembly just because you asked
me nicely.
\stop

\startrobespierre
You continue to misunderstand. I am asking why you have not sent your resources
to the Tuileries itself.
\stop

\startdanton
Oh... well, if that's what you're on about...

\beat{deep breath}

Listen, I know there's been talk about it since that stupid letter from
Brunswick... but, truth is, I don't see the need for putting the boys at risk
when they could be needed for, you know, the army already heading our way.

\brissot and his buddies seem like they have the legality of this forfeiture
business figured out.

So, as far as I can tell, it's only a matter of time.
\stop

\startrobespierre
Is it not your ambition to be seated in a new revolutionary assembly?

Do you not wish to have a say in deciding the future of France's first
republic?
\stop

\startdanton
Is it my "ambition" to give up the influence I have now for the pleasure of
listening to idiots ten hours a day? Fuck no.
\stop

\startrobespierre
You would rather let men like \brissot have such power?
\stop

\startdanton
\beat{chuckles}

Oh, you re-e-eally want that seat for yourself, huh?
\stop

\startrobespierre
You --
\stop

\startdanton
-- Shut up. Call me skeptical of this whole thing you have with \brissot.

I've read your publications, you make up many accusation against the guy, but I
have yet to see you provide any evidence that would hold up in a court.
\stop

\startrobespierre
Evidence?
\stop

\startdanton
Yes. Proof.
\stop

\stage \robespierre hesitates~- he underestimated \danton.

\startrobespierre
A reasonable request.

\beat

Regrettably, I have none to offer.
\stop

\stage \danton is prepared to show him the door.

\startrobespierre
However. Might I suggest an alternative?
\stop

\startdanton
A what?
\stop

\startrobespierre
In the place of evidence, I shall offer you... a prediction.

\beat

Should my prediction come to pass, then you will find in it the evidence that
you seek.

Should my prediction fail, the only cost to you will have been the time you
have already given me.
\stop

\stage Pause.

\startdanton
What's the prediction?
\stop

\startrobespierre
Do we have an arrangement?
\stop

\startdanton
What's the prediction?
\stop

\stage Pause.

\startrobespierre
I predict that when the time comes for the assembly to act, \brissot will fail
to depose the king.

Instead, I predict that the assembly will choose to merely {\emph suspend} the
king.
\stop

\stage \danton laughs.

\startdanton
Bullshit! No way he does that. The public would tear the guy to shreds!
\stop

\startrobespierre
Or... he hesitates to destroy the very throne that he wishes either for himself
or someone that he trusts to carry out his agenda.
\stop

\stage \danton scoffs.

\startrobespierre
Would such an outcome serve as sufficient evidence?
\stop

\stage Pause.

\startdanton
I'm telling you: no man in their right mind would suspend the king.

That's the exact same slap on the wrist that the royalists gave him for trying
to flee to fucking Austria!
\stop

\startrobespierre
Then we have an arrangement.
\stop

\stage Pause.

\stage \danton begins to laugh.

\startdanton
Yeah. I suppose we do...
\stop

\stage The two shake hands.

\stage \robespierre moves to exit, then stops.

\startrobespierre
Do you know what I hate most about him?
\stop

\stage Pause. Silence.

\startrobespierre
I hate how he speaks to people.

\beat

I hate how he pretends to be a philosopher.
\stop

\stage Exit \robespierre.

\stage \danton is a bit stunned. He snaps from his stupor and takes the
untouched drink.

\stage \danton stands to begin reconciling his thoughts. He concludes by
taking a deep breath through the alcohol's aroma.

\startdanton
Camille, my boy, that friend of yours is not a cuddly kitten.
\stop

\stage \danton downs the drink.

\stopcomponent
