\startcomponent eliza

\scene

\open Winter, 1822. The residence of \vernet[madame]. The muted sound of high
winds are heard.

\stage A knock at the door.

\stage Enter \vernet to answer.

\direction The wind makes things difficult to hear.

\startvernet
What in the lord's name makes my company worth the visit in such cold?

Who are you, dear?
\stop

\stage Pause. The wind muffles a response.

\startvernet
Damn it, dear. I cannot hear you! Step in here so I can make sense of you!
\stop

\stage Enter \eliza.

\direction The wind is muted by the shutting of the door.

\startvernet
Who are you, dear?
\stop

\starteliza
Pardon me, madame. Am I correct in assuming this to be the \vernet residence~-
are you the widow of the estate?
\stop

\startvernet
Darling, you ignored my question. Who are you?
\stop

\starteliza
I suppose that's a yes. And - yes, excuse my imposture.

I have been told that you once knew my father.

That you had provided him shelter here in the year, uh, seventeen ninety-three?

My name is \eliza[full].
\stop

\stage \vernet is suddenly swept with emotion! She embraces \eliza~- who
doesn't know how to react.

\startvernet
OH, MY SWEET BABY CHILD!

\beat

My lord! MY LORD! Please! Come in!

\beat

Allow me to fetch us some coffee.
\stop

\starteliza
Oh, no thank you. I have noticed that coffee gives me troubled sleep.
\stop

\startvernet
Troubled sleep is often what I'm looking for, darling.
\stop

\stage \vernet wags her hips at \eliza. \vernet chuckles as she whisks herself
away.

\stage Pause. Silence.

\stage \eliza carefully studies the room~- as if searching for a ghost.

\stage Enter \vernet with two cups of coffee.

\startvernet
Dearest! You must tell me how your mother is doing! Is she well?
\stop

\stage \eliza snaps to attention: "no."

\starteliza
Actually, my mother is what brings me here. She... well, she had passed away
this last September.
\stop

\stage \vernet is stunned.

\starteliza
Before she had passed, she had told me that some of my father's writings had
been kept with you, and~- well, I suppose that she wished for me to continue
her work in ensuring their publication.
\stop

\stage Pause.

\startvernet
Your mother never told you about your father?
\stop

\starteliza
Hm? No, she most certainly has. She would talk about his philosophy all the
time.

\beat{chuckles}

In fact, I would often overhear her talking to him like he's her imaginary
friend.

When I had asked her about it, she told me that he helps her organize her
thoughts.
\stop

\startvernet
She only spoke of his philosophy?
\stop

\starteliza
Yes. About his beliefs of liberty and equality, the virtues of~--
\stop

\startvernet
-- NONESENSE! Darling, have you not learned of his fate?
\stop

\starteliza
Well, of course I know about the revolution! I know of his participation! I
know he succumbed to the guillotine like the rest of the Girondins at the hands
of villainous men.
\stop

\startvernet
Oh, child. Oh, oh, my sweet child. This is unacceptable. This {\emph must} be
remedied!
\stop

\startsplit

\starteliza
I don't --?
\stop

\split

\startvernet
-- FEAR NOT! For I, darling, am now an angel of your destiny!
\stop

\stopsplit

\stage \eliza stands, uncomfortable.

\starteliza
My apologies, madame. But I am mostly interested in the documents.
\stop

\stage \vernet swats at the notion like it's a pest.

\starteliza
Did I offend you? Do you not have them?
\stop

\stage Pause. \vernet smiles.

\startvernet
I do have the letters.

\beat

However, I pray that you are in no rush to enter back out into this biting
weather.

I'll offer you a room for the night and a meal in the morning.

Afterward, you may compensate me with your attention to my tale.
\stop

\starteliza
That. Yes, that sounds excellent.
\stop

\stage \vernet gives \eliza another unanswered embrace.

\startvernet
My apologies for having raised my voice, dear.

\beat

Even if I am an angel of fate and destiny.
\stop

\starteliza
I suppose I've grown used to loud passions.
\stop

\startvernet
Now. Allow me to fetch those letters.
\stop

\stage Exit \vernet~- taking her coffee with her.

\stage \eliza sits back down and eyes her own coffee.

\stage Pause. \eliza makes her decision and partakes.

\stage Enter \vernet~- carrying a trunk with both hands while balancing her
coffee on top.

\stage \vernet sets down the trunk.

\starteliza
He wrote a trunk's worth of dialogue under your care?
\stop

\startvernet
No, dear. This box is where I keep all my letters.
\stop

\stage \eliza opens the trunk.

\starteliza
Oh...
\stop

\stage \vernet smirks.

\starteliza
This is going to be a migraine to sort.
\stop

\stage \eliza gets comfortable and begins. \vernet then joins her.

\stage Pause while the two establish their workflow.

\startvernet
Might I ask. Why choose such terrible weather to pursue this in?
\stop

\starteliza
There are fewer arrests during bad weather. It makes walking the streets safer
for me.
\stop

\stage \vernet takes a deep breath and shakes her head.

\startvernet
We French clearly cannot handle politics.
\stop

\starteliza
All of our faults are rooted in a failure of education.
\stop

\startvernet
Maybe so, dear. But there is certainly some education that is better left
forgotten.
\stop

\starteliza
What?
\stop

\stage \vernet stops sorting.

\startvernet
Child, I have lived to see France go from a monarchy, to a constitutional
monarchy, to a republic, followed by another republic, followed by a consulate,
then an empire, then a monarchy, then back to an empire, until now another
monarchy.

\beat

I have been sampled all of Roman history because men cannot let go of their
obsession with role-playing as heroes of the past.
\stop

\starteliza
But, surely, this does not imply that it should be forgotten.

What you describe is a fault of our institutions, their structure, and their
hierarchies.
\stop

\startvernet
\beat{in a shrill voice}

Humans shape their tools, which in turn shape themselves!
\stop

\stage \eliza is bewildered. She stops sorting.

\starteliza
What was that voice?
\stop

\stage \vernet laughs.

\startvernet
That is how your father sounded, dear!

\beat{imitating \condorcet}

"Truth has prevailed! Humanity is saved!"
\stop

\starteliza
That couldn't possibly be how he sounded, you jest.
\stop

\startvernet
I'm sorry, dear, but I am afraid it's true.
\stop

\stage \eliza puzzles on something.

\stage \eliza returns to sorting.

\startvernet
Oh, how this weather reminds me of the day I met him.

\beat{with flourish}

A summer storm in seventeen ninety-three, with winds equal to that which we
hear against these very walls.

When against this very door was the knock that had forever shaped my fate and
my role in yours.

\beat

I open the door to be presented with a soaked physician~- your uncle, \cabanis.

\beat

His breath was heavy, as if he had just ran across the face of Paris in a hurry
to enjoy the sight of my own.

\beat

And he says to me, "I am a friend of Pinel and Boyer, do you have a room
available for rent?"

And I watch as the man patiently waited for my response while Poseidon
ferociously wept on him.

I say to him, "come in, dear, I cannot hear you!"
\stop

\stage \eliza chuckles.

\startvernet
When he steps inside, he restates his question.

I say, "both of my rooms are occupied."

But your uncle was not dissuaded by this.

By the heaviness of his breath, the pain in his eye~- a sudden twist of his
mouth, as if trying to hide an offensive inquiry that was stuck between his
teeth.

I see in him a man in true need of shelter.

And I say this to him: "my dear, this weather is terrible, if you need my
shelter, you may have it."

\beat

He then apologies! I know not what for!
\stop

\stage \eliza chuckles.

\startvernet
He tells me that he has learned that his brother has been named an enemy of the
revolution - a threat to public safety.

And that he had seen proof that his brother's sentence was to be the
guillotine.

\beat

Then, with great sorrow, he asks me if I might shelter this outlaw.

\beat

He says to me that doing so shall make me an outlaw.

That if he is ever discovered, that I too shall meet God through a blade.
\stop

\stage Pause.

\startvernet
I say to him, "is he virtuous?"
\stop

\stage Pause.

\startvernet
"The most virtuous man I have ever known," is his reply.

\beat

But, dear, such times would make you believe Paris had descended into hell.

Under such circumstances, these words should certainly make me anxious.

But you see, it was not the words that touched me, it was the moment in his
eyes when he spoke it.

Your uncle certainly felt the cold flames of hell when he had first presented
himself to me, but in this glance I see in him... a moment of respite.

\beat

I say, "let him come."
\stop

\stage Pause.

\startvernet
That evening, your father steps through my door, unaware that he had just
stepped into a cell that he shall not step out of for the next nine months.

He comes in soaked by the storm and hardly lifts his head to greet me.

When I looked in his eyes, they were always somewhere else, rarely did he ever
make eye contact with me.

When I spoke to him, he would reply succinctly with a cold, unfeeling
intonation.

Hardly anything like the heroic outlaw I had hoped for.

\beat

I tell him that I house two other men with me, each belonging to the two
bedrooms available for rent.

But I may yet shelter him in the third bedroom that I shall fashion for him.

\beat

"As long as I have access to a light," is all he replies.
\stop

\starteliza
Who were the other two?
\stop

\startvernet
First was my cousin, \sarret.

He stayed with me while he was studying to become a geometer.

The other was a former professor of mathematics and a montagnard by creed.

His name was \marcoz.

\beat

The former knew your father by reputation.

The latter recognized him on sight.
\stop

\starteliza
A montagnard?
\stop

\startvernet
Oh, yes. But he was not a worry. I had a chat with him.

His connection to the party was what allowed your father to receive news of
the convention's doings while he remained hidden beneath its shadow.
\stop

\starteliza
He swore himself to espionage against his party? Why?
\stop

\stage \vernet smirks.

\startvernet
Because he chose me, dear.
\stop

\starteliza
Oh.
\stop

\startvernet
You should consider becoming a widow, darling. It's glamorous.
\stop

\stage \eliza chuckles.

\starteliza
No.
\stop

\stage \vernet laughs.

\startvernet
Nonetheless, your father didn't mind the arrangement with \marcoz.

In fact, he said, it was valuable to learn what decisions were being made at
the \convention.

\beat

I ask him: "what use is there in learning these decisions when they only
frighten and anger you?"

\beat

Because they're ANIMALS," he cries!

O, how I shook my head at him!
\stop

\stage Pause.

\startvernet
From here, things only grew worse.

His routine had reduced to an obsession with his writing.

Truly obsessed!

\beat

In one hand, he would write desperate pleas to the public.

To explain, in as much precision as that imagination of his could derive, all
the reasons that justify him for the decisions he had made while in the public
eye.

\beat

In the other hand, he would write the most disgusting libels.

Accusations laid against men like \robespierre that they were nothing but
ignorant, evil royalists who sought only the destruction of France and her
people.
\stop

\stage Pause.

\starteliza
That's not what I've been seeing so far.

Well, there were a few pages that might be part of what you discuss, but the
tone of many of these documents is~--
\stop

\startvernet
-- Hopeful.
\stop

\stage \eliza nods.

\startvernet
What you observe is the writing of a different man than the one I describe.

The man I speak of was nothing but a frightened child, and I had become like a
second mother to him.

\beat

And lo!

How terrible his cries when \brissot was brought to the guillotine!

O, how he wept into my bosom!

\beat

Such ferocity of his tears had convinced me that no man could ever hope of
recovering from such trauma.

This was all but confirmed by the nights that followed, when your father's deep
malaise of the soul had invited the presence of a demon!

In these nights, I would hear of his encounters with its visitation: I would
hear your father beg from this entity, shout at this entity, lecture to this
entity.

Yet, worst of all, was that this demon would always refuse to leave him~- even
at the embarrassing sight of his lowly sobs and pleas.
\stop

\stage Pause.

\startvernet
Then, one night I receive another knock on this door.

When before me stood a hooded rogue who, in a hush, introduced herself to be
\sophie[condorcet].

Of course, I welcome her.

\beat

When she revealed her face to me, she had a fierce look in her eyes, as if she
was silently, motionlessly rehearsing lines from a play.

"Something is wrong with my husband," she says.

\beat

"Alas! It is true!" I cry!

"A demon has taken residence in his soul, and I fear that no priest can be
consulted for its removal!"

\beat

Your mother shakes her head at this.

"It is an {\emph affliction} of the mind," she replies!

\beat

Naturally, I ask: "my dear, what remedy could possibly cure such an
affliction?"
\stop

\starteliza
Reason.
\stop

\stage \vernet gasps!

\startvernet
You know of this! Your mother has spoken to you about this!
\stop

\starteliza
Perhaps not in the detail you intend to provide.
\stop

\startvernet
No, please, dear. I must hear this.
\stop

\starteliza
Hear what? I don't know what transpired~- I was not in the room when this
happened.
\stop

\startvernet
I must know the words that she spoke to your father that night!

What was it that she meant by reason?
\stop

\starteliza
But you were there to witness this yourself, were you not?
\stop

\startvernet
Oh, my dear! Oh, darling! Oh!

I was not! I was NOT!

\beat

DAMN ME! A curse upon my manners!

Why did I not spy on them!?

\beat

Thus, you see, darling! You are the only chance I have in completing this
missing link!
\stop

\starteliza
My apologies, but I can't be of help to you.

My mother didn't share with me what she said to my father.
\stop

\startvernet
Nonsense! What nonsense! Why would she keep such a secret to her grave?
\stop

\stage \eliza stops sorting.

\starteliza
Nonsense?
\stop

\startvernet
She must have told you of this story since you certainly foresaw that reason
had been my only clue.
\stop

\starteliza
Well, yes, this is true~- but I must insist that my perspective will only lead
to your disappointment.

It will fail to complete your tale.
\stop

\startvernet
Nonsense! Speak up, darling! The lord would not have sent you to me had he not
deemed this essential to our fate!
\stop

\starteliza
Very well.

\beat

During the last few weeks that I had my mother, I asked her by the bedside what
she believed the most difficult moment of her life had been.

\beat

She told me that, by all rights, this question should be impossible for anyone
to answer.

Or, at the least, that it should be impossible for anyone to answer honestly.

Yet, that despite this, she did know such moment.

And she knew I would know this answer to be the truth because of the history
she and I had shared together.

\beat

And she spoke to me of that night.

But she said that her most difficult moment was not the discussion itself, but
in the moments just before she entered his chamber.

She had spent all of the previous six weeks contemplating how this moment would
transpire, considered every possible path that this conversation could take,
plotting contingencies for every unexpected thought her husband could throw her
direction.

Yet, no amount of preparation can deny one particular fact: that in all of
these possible worlds, she was certain that in all of them, sympathy shall make
her suffer more than anything she had ever suffered before.

\beat

She then hesitated. I notice that my mother's eyes had begun to swell with
tears.

And it was then when her promise to me had come true: I was convinced that
there was no greater fear my mother had ever faced.
\stop

\startvernet
Surely, this would not satiate your burning curiosity!?
\stop

\starteliza
Admittedly, I did think twice about asking what was discussed...

Then, quite brazenly, I did.

\beat

She told me that on his part, she didn't listen to a word of it.

But for her part, well, she insisted that if she merely repeated the words that
they would mean nothing to me...

That no man or woman on Earth would understand the meaning of those words...

Except for my father.
\stop

\startvernet
Ugh! That is no excuse!
\stop

\starteliza
It was what she said.
\stop

\startvernet
No! This makes no rightly sense!

No woman would dare keep their greatest achievement forever hidden from
history!
\stop

\starteliza
I apologize if this upsets you.
\stop

\startvernet
UPSETS ME? My very own guest is refusing to share with me the absent truth that
has plagued me for more than twenty years!
\stop

\starteliza
Madame! I don't understand why you think so highly of the particulars!

What makes this matter so severe?
\stop

\stage Pause.

\startvernet
Whatever happened in that room, whatever emotions were revealed, whatever
buried truths were surfaced, whatever decisions made, or morals sealed:
something about your father had forever changed that night.

For the next morning, I witnessed him transform from the sniveling, frightened
child I once knew...
\stop

\stage Pause. \vernet is emotional.

\startvernet
... into the single most hopeful man I have ever known.

\beat

Dear... I must know what was said that night because I must know what words
could possibly have such power over the human spirit.
\stop

\stage Pause.

\starteliza
I suppose the words always depend on the spirit you're speaking to.
\stop

\startvernet
But what sense can there be in that?
\stop

\starteliza
Sense of what, madame?
\stop

\startvernet
The sense of...

A sense of...
\stop

\stage Pause.

\startvernet
Oh! You know, dear! You seem just awfully preoccupied by these papers.
\stop

\stage \eliza is dumbfounded by the non sequitur.

\starteliza
I suppose you could say that. Although, it would help to have assistance.
\stop

\startvernet
I'm sure it would, darling, but I'm afraid that I cannot be of help.

I think it's about time I retire to my chambers.
\stop

\starteliza
Have I offended you?
\stop

\startvernet
Yes, I am offended, dear.

But, fret not, it wasn't you who offended me.

It was merely... your words, I suppose.
\stop

\stage \vernet grabs her coffee. Her hands visibly shaking.

\startvernet
Think nothing of it~- I shall be restored by the morning!
\stop

\stage Exit \vernet.

\stage Pause.

\stage \eliza picks up a document and begins reading.

\stage \eliza takes a deep breath, then returns to sorting.

\stopcomponent
