\startcomponent et_tu_brissot

\scene

\open Night. Outside the \jacobin club. The scene is lit by a streetlamp.

\stage Enter \condorcet and \brissot.

\startcondorcet
War, \brissot? Truly, you wish for war? I-I thought -- y-you've never mentioned
that you supported --
\stop

\startbrissot
-- Calm yourself, my friend. Calm.

You're becoming that mad sheep again!

Recall what you've said to me: that we are all...
\stop

\stage \condorcet recoils, takes a breath, then nods.

\startcondorcet
That we are all born with the ability to receive sensations; to perceive them
and to distinguish them.
\stop

\startbrissot
To remember, recognize --
\stop

\startsplit

\startbrissot[continued]
-- and combine them.
\stop

\split

\startcondorcet
-- and combine them.
\stop

\stopsplit

\startcondorcet[continued]
To compare their combinations; identify what is common and what is not.

To represent their discoveries with signs and symbols.
\stop

\startbrissot
\(chuckles)
And to link these symbols to the eternal chain of human destiny!
\stop

\startcondorcet
Taking us one step forward, to a destination infinitely far away!

\(beat)
You're right, \brissot. My apologies for becoming angry. I must refute your
arguments with reason!
\stop

\startbrissot
I'm afraid now is not the most prudent time.

And debates are tiresome, are they not?

We only have energy for so many.
\stop

\startcondorcet
That's true. We shall arrange a time at \sophie's salon.
\stop

\startbrissot
This sounds excellent.

\(beat)
In the meanwhile, knowing you, I suspect you may have anxieties that linger, so
allow me to help address those.

But this is sensitive gossip, so avoid letting others know I told you these
things.
\stop

\stage \condorcet is a bit off-put.

\startbrissot[continued]
It's true.
\stop

\startcondorcet
Well, I suppose I should respect your interests. Alright, I won't share this.
\stop

\startbrissot
The truth is, by having the national assembly declare support for war with
Austria, we can force king \louis to take a side in the conflict.

\(beat)
If he declines it, he will be declaring his support for Austria, and this would
be treason.

If he supports it, he will be forced to denounce his allies - making him weak.

\(beat)
And it's true that war is already inevitable. I have spoken to general
Dumouriez - he confirms it, but insists that it will only be a short war, about
a few months.

Thus, if the war is already coming, it makes sense to strike with advantage and
utilize its strategic value against the royals.
\stop

\startcondorcet
Ah, I see your reasoning...
\stop

\startbrissot
And I know you. I know what you think of when you think of war.

But this - this is not such a war! We will be liberators! We will enter their
kingdoms, fight any who resist liberty, and welcome the oppressed as allies!

France shall be the model for the future of mankind!
\stop

\startcondorcet
But doesn't this assume --
\stop

\startbrissot
-- Ah, we shall save this for our meeting.
\stop

\startcondorcet
Right.
\stop

\startbrissot
Oh, in addition... Make sure to be cautious of \robespierre...
\stop

\startcondorcet
What? Why must I be cautious? His argument was effective and was delivered with
decorum.

And you yourself had welcomed him to do just that at the end of your speech.
\stop

\startbrissot
Yeah, well, his argument was just a little... too effective...
\stop

\startcondorcet
Right, well, considering I've given my speech, I think I shall~--
\stop

\startbrissot
-- Wait! There's a crowd coming!
\stop

\direction The \sansculottes are heard from a distance.

\startcondorcet
Oh, dear, I think it would be best if we~--
\stop

\startmarat[offstage]
-- If the \jacobins reject our petition, we shall have nine hundred heads
tonight! SO HELP ME GOD!
\stop

\startsansculottes[offstage]
Death to intriguers! Death to schemers!
\stop

\startmarat[offstage]
DEATH TO THE CABAL!
\stop

\stage Enter \danton, carrying a few documents.

\startdanton
{\WORDS\brissot}! Good to see you this evening, Monsieur patriot!
\stop

\stage \danton gives \brissot a hardy handshake. \brissot handles him
excellently.

\startbrissot
\danton, what brings you and the, uh...

\(looks over \danton's shoulder)
... the other Cordeliers to the \jacobin?
\stop

\startdanton
To start an insurrection, of course!
\stop

\startbrissot
So, uh, so soon?
\stop

\stage \danton gives a bellowing laugh. \brissot feigns joining him.

\startdanton
I'm just fucking with you, brother! I'll tell you all about it inside, heh?
\stop

\startcondorcet
I'm actually a bit fatigued... I think~--
\stop

\startdanton
-- Ah! And how's the philosopher doing?
\stop

\startcondorcet
Oh, I have been excellent, t-thank you.
\stop

\startdanton
Glad to hear it! Not letting all the thinking keep you from satisfying that
adorable wife of yours, is it?
\stop

\startcondorcet
Oh.
\stop

\startbrissot
\danton! I think the club is ready to exit recess.
\stop

\startdanton
They must have heard the herald of Poseidon.
\stop

\stopcomponent
