\startcomponent et_toi_robespierre

\scene

\stage \robespierre stands and produces a roman salute.

\stage The shouting halts.

\stage \robespierre[theincorruptible] takes the tribune.

\stage Pause. Silence.

\startrobespierre
Silence? What is this I hear?
\stop

\stage Pause. Silence.

\startrobespierre
In the stories of the ancient greeks, we are told the tale of poor Sisyphus:
the king who wished to defy death, and, for his transgression, was condemned by
the gods to forever roll a large boulder up a mountain.

And each day, king Sisyphus would eventually break from exhaustion, causing the
boulder to roll down to the plain below.

And yet, each morning: he must return to his duty.

Each day. Breaking. From. Exhaustion.
\stop

\stage Pause. Silence.

\startrobespierre
It has been said that many in Paris feel as if they must suffer this fate.

Indeed, I too sympathize with poor, king Sisyphus.

But I feel like no king, and I feel this way for a reason, do I not?

\(beat)
Alone, I am not Sisyphus, but a petty montagnard.

A wretched man with no understanding but earth and gravity~- destined to
scavenge the curvature beneath my feet in search of virtue.

\(beat)
Each day. With my eyes gouged and ears punctured. I break myself on the
mountain.

I do so because my labor is my truth.

My labor is both the will of the people, and the will of the supreme being.
\stop

\stage Pause. Silence.

\startrobespierre
I shall remind you, citizens, that we... We are to be better than kings. We are
to be a {\emph nation}!
\stop

\starttheincorruptible
Alone. I am the montagnard.

But together. WE. SHALL BE. THE MOUNTAIN.
\stop

\stage \themountain riotously cheer.

\stage \robespierre pictures himself as a roman senator~- a childhood dream has
come true.

% TODO: ET TOI!? {\WORDS\robespierre}?

\stopcomponent
