\startcomponent execution_aux

\scene

\open The Place de la R\'evolution. All of \paris is witness. The \executioner
waits by a primed guillotine.

\stage Enter two \patriots with \louis shackled between them.

% TODO: \stage \louis suddenly squares up with the audience, and in an unimpressive
% voice speaks:

% \direction The drum roll stops.

\stage Enter \voltaire.

\startvoltaire
Traveling across northern France, a man beset by the weight of his words, the
weight of his errors, and the weight of his uncertainty watches eastward
through the window of his carriage into the pitch of night.

His wife sits beside him as she pleads with a young man he has never met.

She speaks to him of her fear of the public: the violence that she has seen,
the hatred she has heard, and insists to the young man how she only wishes for
the safety of her children.

\(beat)
But then, something stops her. Choked up by a memory of a time long ago.

In a small, broken voice she recalls her hamlet back in Versailles - the
quaint, rustic little retreat where she would escape the banal politicking of
the palace.

Despite most of the hamlet's beauty being handled by caretakers - she notes
that there was one particular cottage, the smallest of the lot, that was hers
and hers alone.

\(beat)
She remembers a particular morning when she was tending the garden outside this
cottage, that she had looked up to see, across the pond, the first glimpse of
sunrise.

Her voice paints the scene with bright, sentimental colors; delivered with a
sincerity that is not lost on the enchanted young man.

\(beat)
Yet, as the beset man listens, he begins to cry.

He hides his shame by directing his attention to the void outside his window,
furrowing his brow as if to search the darkness - desperate to reclaim
something that has been lost.
\stop

\stage Sharp breaths are heard from the scaffold.

\startvoltaire
"Tomorrow," he thinks... "tomorrow, I shall watch the sunrise."
\stop

\stage The \executioner yanks the rope.

\stage Silence.

\stage Exit \voltaire.

\stopcomponent
