\startcomponent execution

\scene

\open The Place de la R\'evolution. All of \paris is witness. The \executioner
waits by a primed guillotine.

\stage Enter two \patriots with \louis shackled between them.

\stage \louis[thelast] suddenly squares up to the audience:

\startlouis
I die innocent of the crimes that have been charged against me!

I accept my fate, my seat in heaven, and forgive those who are responsible for
my mortal body!
\stop

\direction A drum roll begins.

\startlouis
My only wish is that my blood not be spilt back onto France!
\stop

\stage \louis is shoved onto the scaffold.

\stage Pause. The rustling of leather straps, then the creaking of wooden
joints.

\direction The drum roll ends.

\stage Enter \voltaire.

\startvoltaire
Traveling across northern France, a man beset by the weight of his words, the
weight of his errors, and the weight of his uncertainty watches eastward
through the window of his carriage into the pitch of night.

His wife sits beside him as she pleads with some deputy he has never met.

She speaks to the young man with the kind of sweet sincerity you would expect
from a noble who is begging for their life.

She confesses her fear of the public: the violence that she has seen, the
hatred she has heard, and insists to the young man how she only wishes for the
safety of her children.

\beat

But then, something stops her. Choked up by a memory of a time long ago.

In a small, broken voice she recalls her hamlet back in Versailles - the
quaint, rustic little retreat where she would escape the banal politicking of
the palace.

Despite most of the hamlet's beauty being handled by caretakers - she notes
that there was one particular cottage, the smallest of the lot, that was hers
and hers alone.

\beat

She remembers a particular morning when she was tending the garden outside this
cottage, that she had looked up to see, across the pond, the first glimpse of
sunrise.

Her voice paints the scene with bright, sentimental colors; delivered with a
longing that is not lost on the enchanted young man.

\beat

Yet, as the beset man listens, he begins to cry.

He hides his shame by directing his attention to the void outside his window,
furrowing his brow as if to search the darkness - desperate to reclaim
something that has been lost.
\stop

\stage Sharp breaths are heard from the scaffold.

\startvoltaire
"Tomorrow," he thinks... "tomorrow, I shall watch the sunrise."
\stop

\stage The \executioner releases the blade.

\stage Silence.

\stage Exit \voltaire.

\stopcomponent
