\startcomponent extraordinary_commission

\scene

\open Fall 1792. Committee room of the Tuileries~- lit by candlelight.
\condorcet writes alone at a distinguished desk amongst seating for twenty
others. On the walls are broadsheets with headlines striking fear of war,
famine, and a possible insurrection.

\stage \condorcet lifts his writing to judge it, then stares for a moment at
one of the broadsheets on the wall.

\stage \condorcet draws a large slash through the page and tosses it onto a
growing heap of paper.

\stage He tries again.

\stage Enter \brissot, carrying a few rolled up broadsheets.

\startbrissot
Do you ever stop writing?
\stop

\startcondorcet
Only when I start thinking.
\stop

\stage \brissot laughs.

\startbrissot
But not when you begin speaking?

Is that why you never make eye contact while in conversation?

Do your eyes instinctively seek a parchment to write on instead?

\beat

And not even when you begin sleeping?

Is that why you instead prefer to sleep during our briefings?
\stop

\startcondorcet
Personally, \brissot.

\beat{grins}

I find all of those things to be tedious.
\stop

\stage \brissot laughs.

\startbrissot
Despite your extraordinary honesty, I am convinced that I might never
understand how you think, \condorcet.

I think that's my favorite thing about you.
\stop

\startcondorcet
That... Thank you. I appreciate you saying that.
\stop

\stage \brissot chuckles as he takes a broadsheet and begins attaching it to
the wall.

\direction Headline: "Fanatical Priests Subvert War Effort!"

\startbrissot
Are you preparing your speech for abolishing the throne?
\stop

\startcondorcet
Oh, yes. I'm just establishing my framework for deriving my conclusion.
\stop

\startbrissot
You know, I'm astonished that you've written so much to simply "establish" your
framework.

\beat

\condorcet... I was really hoping you wouldn't do this when I supported you for
president of this commission.
\stop

\direction Next headline: "Royalists in Conspiracy with the Crown!"

\startcondorcet
Nonsense. You supported me because I can be trusted to evaluate the law with
rigor and reason from it possible solutions to public welfare.
\stop

\startbrissot
\condorcet. What are you talking about?
\stop

\stage \condorcet stops writing.

\startcondorcet
What do you mean?
\stop

\startbrissot
What? I'm saying you're not making sense.
\stop

\startcondorcet
I'm asking what specifically does not make sense?
\stop

\stage \brissot is flustered and moves to steal \condorcet's work in progress.

\startbrissot
Oh, my~--

\beat

-- God {\emph damn} it!

\condorcet what is this?
\stop

\stage \condorcet is frozen.

\startbrissot
This is what you've spent the last week doing? Math!?
\stop

\startcondorcet
It's the calculus of probabilities...
\stop

\startbrissot
The {\emph what}? You need to be writing a {\emph speech}! \condorcet! This
must be delivered to the assembly {\emph tomorrow}!

Does that not mean anything to you?
\stop

\startcondorcet
But the speech is trivial! What matters is that our conclusion is correct!
\stop

\startbrissot
Are you not a man of your word?

You promised me that you'd have your speech by the ninth of August, that's what
I fucking told the assembly you'd have it by!
\stop

\startcondorcet
I... I~- I... I'm sorry, I didn't~--
\stop

\startbrissot
-- What, you didn't {\emph know} this? You didn't {\emph know} how much
pressure I'm under to deliver on my promises?
\stop

\startcondorcet
But you cannot predict the speed by which mathematics can be done!

It must be assumed that our predictions are in constant need of incremental
corrections!
\stop

\startbrissot
{\WORDS \condorcet}! What the {\emph fuck} are you talking about!?

THIS ISN'T COMPLICATED!
\stop

\startcondorcet
WHAT!? It is ENORMOUSLY COMPLICATED! \brissot, you started a {\emph fucking}
war!
\stop

\startbrissot
What do you mean I started this war? Wha(t)~--?

\condorcet, {\emph you} voted in favor of it!
\stop

\startcondorcet
You were the one who introduced the notion of it!

The public was in no hurry for war until you told them it was necessary for
their happiness!
\stop

\startbrissot
If you hated it so much, then why did you vote for it?
\stop

\startcondorcet
I don't know.
\stop

\startbrissot
How can you not know!?
\stop

\startcondorcet
I don't know! I~- I~- I struggle to understand it!
\stop

\startbrissot
No wonder you bury yourself in your math. You think this is only about
yourself.
\stop

\stage \brissot sweeps up all of \condorcet's writing, then places them in the
nearby fireplace, and burns them.

\stage \condorcet allows this to happen.

\stage \brissot procures a fresh page and places it in front of \condorcet.

\startcondorcet
Okay.
\stop

\stage Pause. \brissot is waiting for \condorcet to begin.

\startcondorcet
Well. Consider the assembly as a set consisting of three non-intersecting
categories: the left, the middle, and the right.

You know which deputies belong to which category based on who and what they
applaud for, with the middle consisting of those who do not applaud at all.
\stop

\stage Pause.

\startcondorcet
You see, the applauses inform us about the preferences of their votes, allowing
us to apply social arithmetic to determine an optimal outcome for public
welfare.
\stop

\stage Pause.

\startcondorcet
To begin, we consider the observation that deputies belonging to the right do
not change their beliefs based on reason.

We know this because these deputies do not applaud whenever improvements are
made to their existing beliefs by deputies belonging to the left.

Therefore, if we desire the outcome of a pairwise-majority decision to improve
public welfare, then our only option is to propose a solution that can win
support from the middle.

So, I wonder, what then should I propose?
\stop

\stage Pause.

\startcondorcet
\beat{stands proudly}

A {\emph compromise}!
\stop

\startbrissot
I'm vetoing all of that from your speech.

\beat

Pick up the damn quill.
\stop

\startcondorcet
You don't have veto power, and I am the sitting president of this commission.
\stop

\startbrissot
How do you even think you're even a deputy in the first place! I was the one
who got you elected at all!
\stop

\startcondorcet
By lying to a whole township of people you have never met about fake votes!
\stop

\startbrissot
Pick up the quill.
\stop

\startcondorcet
Leave me in peace, and I shall.
\stop

\stage Pause.

\stage Exit \brissot.

\stage \condorcet sits down and stares at the page.

\stage Enter \marat and \danton, dressed as \leibniz and \voltaire.

\stopcomponent
