\startcomponent flight

\scene

\open Paris, Summer 1791. \condorcet is reading the daily broadsheets.

\stage Enter \voltaire.

\startvoltaire
Traveling across northern France, a man beset by the weight of the world and
the weight of his fat watches westward through the window of his carriage as
the sun proceeds to set over the farmland just beyond the border of Varennes.

While the six horses ahead of him trot onward, he gazes on the hypnotic
parallax of the wheat fields with a certain wistful satisfaction.

\(beat)
"If only we just had more of these," he thinks to himself.

"France could really use more of these."

\(beat)
The man's eyes soon become bewitched with visions of working the land: he
pictures himself tilling the soil, scattering the seeds, harvesting the yield,
or whatever else those people do all day.

And in this moment, with the smell of manure painting the scene against a
dwindling sunset, he feels an irresistible kinship with such an image.

As if it was by some cruel twist of fate that his heritage had somehow robbed
him of his chance to realize his true self.

\(beat)
Yet, he smiles. Because, for the first time, he allows the limits of his
imagination to finally brush against the boundary of a life as it was always
meant to be in the best of all possible worlds.

\(beat)
A life without words.

A life without errors.

A life without uncertainty.

\(beat)
... well, barring the weather.
\stop

\stage Exit \voltaire.

\direction The singing of La Prise de La Bastille is heard from another room.
An infant begins to cry as a cradle is revealed.

\stage \condorcet puts down the broadsheet and attends to the cradle to lift an
infant into his arms.

\direction \condorcet speaks with a shrill voice.

\startcondorcet
Has the revelry woken you, dear \eliza?

\(beat)
I see why you're upset. You don't yet understand the joys of pride that compels
us to sing for our nation.

As your mother would say, you are only just now beginning to learn the pain of
being separated from your nurse and the joy of being reunited.

\(beat)
I'm happy to oblige.
\stop

\stage Enter \sophie.

\startsophie
\brissot has arrived, and he brought a friend from America!
\stop

\startcondorcet
Is that so? Oo, I hope he brings a letter from Jefferson.
\stop

\startsophie
I'll let them know you'll be down soon.
\stop

\startcondorcet
Thank you.
\stop

\stage Exit \sophie.

\direction The singing comes to an end.

\startcondorcet[continued]
Ah, perhaps a story will bore you back to sleep!

\(beat)
Well, once in the kingdom of New Atlantis, there lived a woman who had since
her childhood been fascinated by the wonders of automata:

machines that bear the likeness and function of animals, flute players, chess
masters, what have you!

Thus, she dedicates her education to becoming an expert in the subject of
mechanics in order to honor the ingenuity of those who had preceded her and
those who had mentored her.

\(beat)
And so it was that on a fateful day, when the king of New Atlantis had
disappeared, the mechanic sprung into action to assuage the fears of her
country.

"Hark, Mademoiselle!" some would retort, "what can you offer us when that which
we have lost was bestowed by divine right?

Any solution would require the claim of divine providence, and this would be
heresy!"

\(beat)
The mechanic replies, "fear not! I do not proclaim myself to be any more divine
than that of yourselves.

My solution, likewise, shall be derived from the very same divinity as that of
the earth and the sky.

For you see, I am a skilled mechanic, and I can promise to build, within a
fortnight, a new king that can fulfill, in equal measure, all the duties of
the court."

\(beat)
"Impossible!" the fearful crowds cry.

"No such machine has ever demonstrated wisdom equal to that of a just monarch."

\(beat)
The mechanic continues, "it is indeed by my very same love for intelligence and
justice that I share with you my proposal.

Consider, good people, that all just kings serve the welfare of their subjects.

Thus, we may represent the will of this plurality by electing a council of
representatives whose role is to prepare well-formed decrees designed to
benefit the public good.

These decrees, thus submitted to the court, shall be signed and sanctified by
the divine right of our new automaton-king and sanctioned by the authority of
our sciences and our passion for the public welfare."

\(beat)
The crowd begins to stir in response to these confounding ideas.

Yet, just as someone begins to object, the mechanic interrupts:

"citizens, I understand there have long existed, among many nations, the rule
of hereditary kings.

But let us read their history and dare to say that they would not have lost
much by following my method.

While it is true that the cost of building our automaton-king will doubtless be
enormous, the maintenance shall be trivial compared to what is demanded by the
current royalty.

Our automaton-king would not impose any threat to liberty, and yet, by
carefully repairing him, he would be eternal, which is even more beautiful than
being hereditary!"
\stop

\stage \condorcet cackles.

\startcondorcet[continued]
Oh, my apologies. I didn't mean for my laughter to wake you.
\stop

\stopcomponent
