\startcomponent jacobins

\scene

\open Jacobin club. The \crowd[jacobins], along with \danton and \brissot, are
in heated deliberation. \robespierre is reviewing a small notebook.

\startanonymous[cue={Jacobin 1}]
We can't blame the assembly for this massacre!

\marat is to blame for corrupting the souls of patriots with violence!
\stop

\startdanton
As far as I can tell, \marat wasn't the one firing muskets on unarmed men,
women, and children!
\stop

\startsplit

\startanonymous[cue={Jacobin 2}]
\danton is right!
\stop

\split
\vskip\bigskipamount

\startanonymous[cue={Jacobin 3}]
Why defend him, \danton? \marat is a fool and a fiend!
\stop

\stopsplit

\startdanton
I'll defend any true patriot, and there is no one more devoted to his country
than \marat!

Sure, the man gets carried away, but what man wouldn't when his passion for the
revolution overwhelms him!
\stop

\startbrissot
Stop this insufferable debate!
\stop

\stage The \crowd[jacobins] deride him.

\startbrissot[continued]
Our only enemy is that fat pig and his ministers!

\louis is corrupt and a corrupter!

It is only to his advantage that we divide ourselves!
\stop

\startanonymous[cue={Jacobin 1}]
You too, \brissot? You would be complicit in \marat's behavior?
\stop

\startbrissot
I wouldn't say I approve of his methods, but --
\stop

\stage The \crowd[jacobins] interrupt him with jeers.

\direction A hostile gathering approaches the doors of the hall.

\startanonymous[cue={Jacobin 2}]
The National Guard! \lafayette and his royalists have come to put a stop to us!
\stop

\direction Loud pounding at the door.

\stage The \crowd[jacobins] recoil in fear. \robespierre is focused.

\direction The door is broken down.

\stage Enter the \crowd[sansculottes] and \marat.

\startmarat
JACOBINS! TO ARMS! JOIN US TO THE ASSEMBLY!
\stop

\startcrowd[cue={Sans-Culottes}]
TO ARMS! TO ARMS!
\stop

\startsplit

\startanonymous[cue={Jacobin 1}]
They would doom us all!
\stop

\blank[big]

\startanonymous[cue={Jacobin 3}]
You would justify our oppressors!
\stop

\split
\vskip2\bigskipamount

\startanonymous[cue={Jacobin 2}]
TO ARMS!
\stop

\stopsplit

\startmarat
COWARDS! The time has come to die for liberty and your mask of patriotism
falls!

YOUR HEADS SHALL JOIN THE SIX HUNDRED THAT FALL TODAY!
\stop

\startbrissot
NO!
\stop

\startcrowd[cue={Sans-Culottes}]
Death to all who read and write!

Death to all who eat with a silver spoon!
\stop

\stage \robespierre closes his notebook.

\startrobespierre
I DEMAND THE FLOOR!
\stop

\stage The room falls to a still.

\stage \robespierre commands the tribune. \marat sneers at him.

\startrobespierre[continued]
Jacobins. You have allowed fear to consume you. You have lost sight of those
who our revolution is fought for.

We fight for the people!

For the poor, the trodden, the weak, the forgotten!

Here the virtuous have arrived to call upon you, and you have the audacity to
tremble.
\stop

\stage \marat drops his sneer: he remains guarded, but is now intrigued.

\startrobespierre[continued]
You would recall, it was only a few months prior that I had risen to the
tribune to warn you that the assembly has become dominated by enemies to the
revolution

And at that time, I was not heeded.

\(beat)
But I had forgiven you then. I had not brought you proof.

However, it was soon to follow that proof did reveal itself to me.

\(beat)
I did not bring this to you at the time due to the confusion it would have
caused, but only days following the return of \louis, I had been approached by
a royalist offering twenty thousand livres in exchange for my support for
martial law.
\stop

\stage The \crowd gasps.

\startanonymous[cue={Jacobin 3}]
Who? Who was it?
\stop

\stage Pause.

\startrobespierre
\barnave. I declined him.
\stop

\stage A chill runs through the room.

\startrobespierre[continued]
People had wondered why I had fought to forbid members of the assembly from
becoming legislators in the next government - despite it being against my own
political interests.

This, citizens. This is why. I made this sacrifice in the fight against the
corrupt.

\(beat)
I have since listened to their plotting, thus once the idea for a petition was
brought to our attention, I alone had warned of the danger it would pose to our
beloved patriots.

\(beat)
Again, I was not heeded. And now, two hundred innocents are dead.
\stop

\startmarat
Five hundred innocents!
\stop

\startrobespierre
\(about to weep)
Too many to bear!
\stop

\stage \robespierre removes his spectacles and rubs his eyes. Deep breath.

\startrobespierre[continued]
But I shall not hesitate! I shall not flinch! I SHALL NOT FAIL YOU AGAIN!

\(beat)
You, \marat! You champion of the people! You warrior of the revolution!

I cannot fail you!
\stop

\stage \marat is moved by this.

\startrobespierre[continued]
You must trust my compassion for our cause!

My compassion for you with patches in your coats!

I alone know who your enemies are.

I know who has taken the other bribes.
\stop

\startcrowd[cue={Sans-Culottes}]
WHO? TELL US WHO!
\stop

\stage \danton becomes agitated.

\startrobespierre
It is not the right time.

But worry not! I hear your plight, I know your struggle. I feel the blood that
emboldens your hate!

\(beat)
So I ask of you. Let me be your sword. Let me guide you to our victory.

We are outnumbered by enemies who yet still lurk in the shadows.

Violence now would compel these beasts to pounce.

Thus, it is our patience that shall be our defense.

\(beat)
I understand that it won't bring you the glory you had expected.

I understand your desire to demonstrate your courage, your heroism, but heed me
- I beg of you.

We must trust in the power of strategy.
\stop

\stage \robespierre takes a deep breath and dons his spectacles.

\startmarat
LONG LIVE {\WORD \robespierre}, THE INCORRUPTIBLE!
\stop

\stage Riotous applause. \brissot and \danton do not contribute.

\startcrowd
LONG LIVE THE INCORRUPTIBLE!
\stop

\stopcomponent
