\startcomponent legislature

\scene

\open The Salle du Man\`ege of the Tuileries. The many \deputies are in the
process of finding their seats. The \deputies make sure to introduce themselves
to their neighbors upon settling on a seat. \brissot is already seated to the
left side of the tribune. Idle discussion fills the room.

\stage Enter \condorcet\ - he shyly surveys the room.

\stage \brissot stands and waves to him. \condorcet notices and joins him.

\stage Enter a \deputy. He appears paralyzed.

\stage Enter \voltaire, who approaches him.

\startvoltaire
Look at this poor sod. Tsk, tsk.

By his outfit, he looks more like a rancher than a politician - certainly not a
Parisian.

I'm guessing he can't be older than twenty-eight.

\(beat)
The lad must be utterly out of his element, like a little schoolboy on his
first day of school.

Which club should he join, do you think?
\stop

\stage The lone \deputy snaps out of his stupor and begins making
friends to the right.

\startvoltaire[continued]
Oh, there he goes! Kudos to him! Shame he picked the wrong side, though.
\stop

\stage \voltaire chuckles to himself.

\stage Exit \voltaire.

\stage The \president of the assembly rises to the tribune and sounds a large
bell.

\startpresident
Silence! Silence on the floor! Take your place!
\stop

\stage The discourse settles down.

\startpresident[continued]
Welcome everyone to the first session of National Legislative Assembly of
France.
\stop

\stage Applause.

\startpresident[continued]
We are honored to have sworn our oath to the wonderful constitution that we
have been blessed with, and now it is time to serve our nation in its honor.
\stop

\direction \condorcet begins ignoring the \president's speech.

\stage Enter \leibniz.

\startleibniz
My word, the acoustics of this hall are abysmal!

This narrow corridor hardly suits its function as a place for legislation!

\(beat)
Observe the geometry of the galleries above us and how they capture and deaden
the sound. What nonsense!

And to make matters worse, observe the positioning of the tribune.

The speaker is positioned as if he speaks only to himself!

\(beat)
Ugh, and the smell. The air is stuffy and oppressive.

Perhaps it would be prudent to have ventilation installed such that --
\stop

\stage From the left, a \jacobin stands.

\startjacobin
I request the floor!
\stop

\stage From the right, a \feuillant stands.

\startfeuillant
I demand decorum! How dare you interrupt Monsieur \president!
\stop

\stage The \jacobin ignores this and storms to the tribune. The \deputies begin
chattering over themselves.

\startjacobin
I motion that the assembly maintain a strict adherence to the constitution in
its deference to the king.

No one shall treat the king in any regard that places him above that of the
president and he shall not be addressed in terms other than the title "king of
the French," as written in the constitution.

No longer should any citizen be required to refer to the king as "sire," or
"majesty."
\stop

\startfeuillant
You dare begin with legislation to insult the king!? Monsieur \president! This
cannot be tolerated!
\stop

\startjacobin
In addition, I motion that the seat of the king be made identical to that of
the president!
\stop

\stage The \deputies begin shouting over one another.

\direction \condorcet tunes out the shouting.

\stage Enter \voltaire.

\startvoltaire
Oh, what did I miss? What exciting, enlightened debates have emboldened our
beloved new government?
\stop

\stopcomponent
