\startcomponent mad_sheep

\scene

\open Outside the \jacobin club.

\stage Enter \condorcet and \brissot.

\startcondorcet
"A people who have won their freedom need war to consolidate it?"

HOW COULD YOU SAY SUCH A THING? You've never - y-you --
\stop

\startbrissot
-- Calm yourself, my friend.

You're becoming that mad sheep again!

Take a breath and recall what you've said to me: that we are all...
\stop

\stage \condorcet recoils, then nods.

\startcondorcet
That we are all born with the ability to receive sensations; to perceive them
and to distinguish them.
\stop

\startbrissot
To remember, recognize --
\stop

\startsplit

\startbrissot
-- and combine them.
\stop

\split

\startcondorcet
-- and combine them.
\stop

\stopsplit

\startcondorcet
To compare their combinations; identify what is common and what is not.

To represent their discoveries with signs and symbols.
\stop

\startbrissot
\(chuckles)
And to link these symbols to the eternal chain of human destiny!
\stop

\startcondorcet
Taking us one step forward to a destination infinitely far away!

\(beat)
You're right, \brissot. My apologies for becoming angry. I must refute your
arguments with reason!
\stop

\startbrissot
I'm afraid now is not the most prudent time.

And debates are tiresome, are they not?

We only have energy for so many.
\stop

\startcondorcet
That's true. We shall arrange a time at \sophie's salon.
\stop

\startbrissot
This sounds excellent.

\(beat)
In the meanwhile, knowing you, I suspect you may have anxieties that linger, so
allow me to help address those.

But this is exceptionally sensitive gossip, so avoid letting others know of
these things.
\stop

\stage \condorcet is a bit off-put.

\startbrissot
It's true.
\stop

\startcondorcet
Well, I suppose I should respect your interests. Alright, I won't share this.
\stop

\startbrissot
The truth is, by having the upcoming assembly declare support for war with
Austria, we can force king \louis to take a side in the conflict.

\(beat)
If he vetoes it, he will be declaring his support for Austria, and this would
be treason that can be leveraged against him using the justice system
established by the constitution.

If he supports it, he will be forced to denounce his allies - cutting him off
from the foreign tyrants he conspires with.

\(beat)
And it's true that war is already inevitable. I spoke with General Dumouriez -
he confirms it.

Not only that, but he also insists that it will only be a short war, about a
few months.

Thus, if the war is already coming, it makes sense to strike with advantage and
utilize its strategic value in seeing \louis deposed.
\stop

\startcondorcet
Oh. I, uh --
\stop

\startbrissot
-- And I know you. I know what you think of when you think of war.

But this - this is not such a war! We will be liberators!

We will enter the kingdoms of tyrants, fight any who resist liberty, and
welcome the oppressed as allies!

France shall be the model for the future of mankind!
\stop

\startcondorcet
But this assumes --
\stop

\startbrissot
-- Ah, we shall save this for our meeting.
\stop

\startcondorcet
Right.
\stop

\startbrissot
Oh, in addition... You should be cautious of \robespierre...
\stop

\startcondorcet
What? Why must I be cautious? His rhetoric was effective and delivered with
decorum.

And you yourself had welcomed him to do just that at the end of your speech.
\stop

\startbrissot
Yeah, well, I found his rhetoric to be a little... too effective...
\stop

\startcondorcet
I agree that his argument had errors, but that is to be expected.

We simply need to explain such errors and make the proper improvements to our
reasoning.
\stop

\startbrissot
That's an interesting way to think about it. I'll certainly meditate on that.
\stop

\startcondorcet
That's very pleasant to hear.

Now: I suppose that I've given my speech, thus I think I shall~--
\stop

\startbrissot
-- Wait! There's a crowd coming!
\stop

\direction The \sansculottes are heard from a distance.

\startcondorcet
Oh, dear, I think it would be best if we~--
\stop

\startmarat[offstage]
-- If the \jacobins reject our petition, we shall have nine hundred heads
tonight! SO HELP ME GOD!
\stop

\startsansculottes[offstage]
Death to intriguers! Death to schemers!
\stop

\startmarat[offstage]
DEATH TO THE CABAL!
\stop

\stage Enter \danton, carrying a few documents.

\startdanton
{\WORDS\brissot}! Good to see you this evening, monsieur patriot!
\stop

\stage \danton gives \brissot a hardy handshake.

\startbrissot
\danton, what brings you and the, uh...

\(looks over \danton's shoulder)
... the other Cordeliers to the \jacobin?
\stop

\startdanton
To start an insurrection, of course!
\stop

\startbrissot
So, uh, so soon? The elections to the Legislative Assembly aren't for another
few months.
\stop

\stage \danton gives a bellowing laugh. \brissot feigns joining him.

\startdanton
I'm just fucking with you, brother! I'll tell you all about it inside, heh?
\stop

\startcondorcet
I'm actually a bit fatigued... I think~--
\stop

\startdanton
-- Ah! And how's the philosopher doing?
\stop

\startcondorcet
Oh, I have been excellent, t-thank you.
\stop

\startdanton
Glad to hear it! Not letting all the thinking keep you from satisfying that
adorable wife of yours, is it?
\stop

\startcondorcet
Oh.
\stop

\startbrissot
\danton! I think the club is ready to exit recess.
\stop

\startmarat[offstage]
NEW SANITATION METHODS WILL BE NEEDED WHEN OUR STREETS RUN WITH BLOOD!
\stop

\startdanton
They must have heard the herald of Poseidon.
\stop

\stopcomponent
