\startcomponent on_the_republic

\scene

\open Jacobin club of Paris. \robespierre is sitting alone, meditating. His
eyes are always obscured by a pair of tinted spectacles.

\stage Enter the various \jacobins.

\stage As the \jacobins notice \robespierre, they all rush to confront him.

\startjacobins
\robespierre! \robespierre!
\stop

\startjacobin[number=1]
\robespierre! Are you here to argue for the republic?

With the king having fled, do you support his deposition?

Do you support the republic?
\stop

\startjacobin[number=2]
All of Paris is in suspense! We must know if you support the republic!
\stop

\stage Pause. Silence.

\stage \robespierre[theincorruptible] produces a sly grin.

\startrobespierre
What even is a republic?
\stop

\stage The \jacobins erupt into deliberation.

\stage Enter \brissot, \paine, and \condorcet, followed shortly by the
\president.

\stage The \president takes to the podium and rings a loud bell.

\startpresident
Gentlemen! Gentlemen! Silence!
\stop

\stage The \jacobins settle down.

\stage \robespierre prepares a small notebook - he intends to take notes.

\startbrissot
Monsieur \president, I request the floor!
\stop

\startpresident
\brissot! The floor is yours!
\stop

\stage \brissot takes to the podium. A healthy number of \jacobins applaud.

\startbrissot
Why do we wait for them to strike?

Why are we so unclear about the stakes?

Yesterday, the \jacobins would have been unanimous about the inevitability of
an Austrian army invading France.

This has been true since the day we learned of the plots by the \'emigr\'es.

\(beat)
And in our indecision, we allowed these emigrants to accelerate their schemes
into fruition!

All they had to do was speak to the emperor about the idea of liberty and now
they have raised an army!

\(beat)
And meanwhile, even, rational minds have succumb to prejudices.

If we are to arrive at the truth, we must banish these prejudices and examine,
with tolerance and clarity, the reasoning of both parties.

We must defend against the passions that lead us into disorder.

\(beat)
Gentlemen, patriots are allowed to differ in opinions!

In fact, this is healthy!

Because it is this same conviction we have for ourselves that we must
demonstrate against our enemies!

A nation, after suffering twelve centuries of slavery and having now won its
freedom, must have war to consolidate its victory.
\stop

\stage The \jacobins begin discourse.

\stage Enter \leibniz, then \voltaire.

\startsplit

\startleibniz
War is necessary to consolidate~--?
\stop

\split
\leavevmode

\startvoltaire
-- WHAT? War? What the hell, WAR? He's never mentioned that he wanted war! This
is a betrayal!
\stop

\stopsplit

\startpresident
\(ringing a bell)
Silence! Decorum!
\stop

\stage Exit \leibniz and \voltaire.

\startbrissot
Thank you, Monsieur \president.

\(beat)
A nation needs war so that its strength can be tested.

To ensure that its citizens are worthy of it.

\(beat)
The Americans have braved the crucible of war; and have continued to do so
since the instant they raised the banner of liberty.

Bless heaven that we are only two years late to the call.

Bless the time that it has given you in completing your constitution.

Now, it is a question of defending it.

A question of putting an end to those abroad who seek to destroy it.

Of putting an end to the contemptible intriguers within who would sabotage it.
\stop

\stage The \jacobins applaud.

\startbrissot[continued]
Now, let it be said, I welcome any patriot to defend their perspective to the
contrary.

But I implore them to examine my arguments and provide considered, healthy
refutations.

If I am mistaken, if I am convinced, I shall shock you all when I reveal my
embrace for their position.

\(beat)
But if you instead have your prejudiced misgivings, then speak them now.

For I swear on my life that I shall stand here to destroy your objections until
I drop dead.
\stop

\stage Pause.

\startbrissot[continued]
I thought not.
\stop

\stage \brissot dismounts the podium to great applause - excluding \paine and
\robespierre. \condorcet claps with a weird look on his face.

\stage \robespierre closes his notebook and stands.

\startrobespierre
Monsieur! The floor!
\stop

\startpresident
The floor is yours, \robespierre.
\stop

\stage \robespierre[theincorruptible] takes to the podium.

\stage Pause. Silence.

\startrobespierre
Of any two opinions that split the \jacobins, one has on its side all the ideas
that flatter the imagination, all the brilliant hopes that ignite the spirit
and animate enthusiasm;

the other... is supported only by cold reason and difficult truths.

\(beat)
If I wished to please, I would defend the first.

If instead I wish to survive, I must support the second.

I understand if it shall displease you, but it is for the latter that I declare
myself.
\stop

\stage Pause. Silence.

\startrobespierre[continued]
Shall we make war, or shall we make peace?

Shall we attack our enemies, or shall we await them in our homes?

I believe that these questions fail to appreciate the depths of our conflict's
true character.

\(beat)
If the brilliant and prophetic painting of success truly ended in perfect
harmony and fraternal love for all the people of Europe,

and if this ambition is deemed worthy of something as extreme as declaring war,

then I would agree that Monsieur \brissot has perfectly resolved this debate.

\(beat)
Certainly, I would like as much as anyone to expand the reign of liberty - in
fact, I would gladly indulge in the pleasure of dreaming up its wonders with
you over dinner.

And if I were the master of France's destiny, I would have long ago sent armies
to neighboring tyrants to break the chains of their subjects.

\(beat)
But in the France I see now, I cast an anxious glare.

I ask: who is asking for war? How are they asking for it, in what
circumstances, and why?

\(beat)
I shall remind you, citizens, that all that matters is that we enlighten the
nation of its true interests and reveal to it the interests of its enemies.

So you ask: where are these enemies of the revolution?

They are among yourselves.

\(beat)
The bitter truth about this debate on war is that those who are championing it
are doing so to advance their hidden projects.

An agenda conspired by the internal enemies of liberty long ago.

This, citizens, is a trap.

\(beat)
Consider: who benefits from its outcomes?

Should France win the war, the court, having sanctioned it and who would have
executed it, would be treated as heroes.

This spirit and prestige would then, as it had for Caesar, be used to nullify,
then silence, the general will.

Should France lose, then we will be killed and, of course, the monarchy will be
reinstated.

\(beat)
So, I ask again... who is advocating for war? And why?
\stop

\stage Pause. Silence.

\startrobespierre[continued]
And, for what it's worth, nobody likes an armed missionary.
\stop

\stage \robespierre[theincorruptible] returns to his seat and opens his
notebook.

\startjacobins[cue={A few Jacobins}]
Long live the \robespierre[incorruptible]! Long live the nation!
\stop

\stage Great applause from the \jacobins\ - excluding \paine and \brissot.

\startcondorcet
Monsieur \president, I would like to request the floor.
\stop

\startpresident
Huh?

\(beat)
Oh! Yes, uh, the philosopher wishes to speak!

\condorcet, the floor is yours.
\stop

\stage \condorcet takes to the podium. A handful of \jacobins applaud -
including \paine and \brissot.

\stage \condorcet unfolds his speech and rests it on the tribune.

\stage He is an atrocious orator.

\startcondorcet
T-the citizens of - of France, having demonstrated their courage in claiming
their freedom, and having now demonstrated their calm firmness in - in
contemplating the danger that has just threatened it, no longer require
eloquence in the call to liberty.

It is therefore to their reason alone that we must speak on the security of a
peaceful future that is worthy of an enlightened people.

\(beat)
In the face of recent events, the question that we are all obliged to ask
ourselves is this:

is a king necessary for the establishment of liberty and the protection from
tyranny?

\(beat)
No, gentlemen, it is not.

\(beat)
The friends of royalty would argue that the power of an executive that is
established and limited by written law is far less formidable than that of a
usurped power that has no other limits than those of skill and audacity.

However, these people neglect to consider that a republic~--
\stop

\stage A few \jacobins gasp.

\startcondorcet[continued]
... that a republic, formed from a sound and enlightened constitution, would
see the powers of the army, the finances, and the courts be adequately divided
and distributed between disparate departments and deputies.

Therefore, it would be necessary to adequately divided and distribut(ed) --
oh, I read this part already.

\(clears throat)
Uhm... t-therefore, it would be necessary to have destroyed, corrupted, or
denatured all these powers simultaneously before one could aspire to tyranny.

And since any enlightened constitution would guarantee the freedom of the
press, this threat is therefore impossible for a republic where deputies are
subject to frequent elections.

\(beat)
Consider that any popular tyrant can only act under a mask that hides their
hypocrisy.

Thus, the power ensured by the press to force these ambitious people to walk
with their faces uncovered shall render them all h-harmless.

Let us therefore not allow ourselves to harm the progress of the human species
and the happiness of our nation out of fear for an imaginary danger.
\stop

\stage Polite applause from the \jacobins\ - excluding \brissot and
\robespierre.

\startcondorcet[continued]
No longer do we live in a time when laws can only be upheld through divine
intervention.

No longer are we prejudiced by the impious superstitions that suppose that any
man can speak on behalf of a supreme intelligence so long as he wears the
proper attire.

\(beat)
Yet, those friends of royalty, those who remember the events of history, but
who do not know history, remain frightened by the corruption of ancient
republics.

But let them examine these republics, and they will always see a division
between those who are sovereigns and those who are subjects.

Then they will understand the true forces that corrupt the privileged and
seduce the impoverished, and it will be clear that these forces cannot exist in
a nation where equality is made universal to all people who inhabit it - no
matter their race, their wealth, or - or their gender.
\stop

\stage The \jacobins are confused by this.

\startcondorcet[continued]
It is therefore not the tyranny of an individual that we must fear, but the
most odious and sinister tyrannies that exist when one multitude rules over
another through abuses of law and finance.

\(beat)
But is such abuse what the friends of liberty demand? Those who only accept
reason as the sole master of their fate?

I ask you: at whose expense could we satisfy the greed of our leaders?

What conquered provinces could a French general strip to buy our votes?

Would such an ambitious man propose to us, as he had to the Athenians, to levy
tributes on our neighbors in order to fund new temples and festivals?

Will he promise our soldiers, as he had to the Romans, the pillage of Spain or
Syria?

No, doubtless!

\(beat)
I ask you: how could we feed the greed of the few in a nation that refuses to
exploit their allies?

In a nation that refuses the conquests of war?

In a nation that refuses to be subjugated and built by citizens who refuse to
subjugate others?

\(beat)
I answer you, my friends, that it is because we refuse to be a kingly people
that we will remain a free people.
\stop

\stage \condorcet folds up his speech.

\stage Applause from the \jacobins\ - excluding \brissot and \robespierre\ -
builds from trepidation to enthusiasm.

\stage \condorcet dismounts the podium.

\stage The \jacobins begin deliberating between themselves.

\startpresident
Any further motions?
\stop

\startjacobins
Nay.
\stop

\startpresident
Recess!
\stop

\stage The \jacobins begin deliberating more liberally.

\startcondorcet
\brissot, do you mind if we discussed outside?
\stop

\startbrissot
Ah, I think I understand. \paine[informal], are you content to wait here?
\stop

\stage \paine nods.

\stage Exit \condorcet and \brissot.

\stopcomponent
