\startcomponent on_the_war

\scene

\open Jacobin club of Paris. \robespierre is sitting alone, meditating. His
eyes are always obscured by a pair of tinted spectacles.

\stage Enter the various \jacobins.

\stage As the \jacobins notice \robespierre, they all rush to confront him.

\startjacobins
\robespierre! \robespierre!
\stop

\startjacobin[number=1]
\robespierre! Are you here to argue for the republic?

With the king having fled, do you support his deposition?

Do you support the republic?
\stop

\startjacobin[number=2]
All of Paris is in suspense! We must know if you support the republic!
\stop

\stage \robespierre[theincorruptible] produces a sly grin.

\startrobespierre
What even is a republic?
\stop

\stage The \jacobins erupt into deliberation.

\stage Enter \brissot and \condorcet, followed shortly by the \president.

\stage The \president takes to his chair and rings a loud bell.

\startpresident
Gentlemen! Gentlemen! Silence!
\stop

\stage The \jacobins settle down.

\stage \robespierre prepares a small notebook and pencil - he intends to take
notes.

\startbrissot
Monsieur \president, I request the floor!
\stop

\startpresident
\brissot! The floor is yours!
\stop

\stage \brissot[thepatriot] takes to the tribune.

\stage Half the \jacobins applaud, the other half jeer.

\startbrissot
Yesterday, the \jacobins would have held unanimously that the invasion of
foreign tyrants was inevitable.

This has been true since the day we learned of the plots by the emigrants!

Our sudden indecision allows these traitors to accelerate their schemes and
bolster their treacherous ambitions.

All they would need is to speak the word "republic," and kings and emperors
everywhere would raise an army to our borders on all fronts!

\(beat)
The lives of innocent patriots have been threatened, and meanwhile, rational
minds have succumb... to prejudices.
\stop

\stage The \jacobins begin shouting.

\startsplit

\startjacobin[number=3]
Hypocrite!
\stop

\split

\startjacobin[number=4]
\brissot is a rogue; his soul is as black as the negroes!
\stop

\stopsplit

\startjacobin[number=3, extension=continued]
GO WORK THE FIELDS, "PATRIOT!"
\stop

\startbrissot
How dare you! HOW DARE YOU! GO TO HELL, YOU WHORESONS!
\stop

\startpresident
Silence! Decorum!
\stop

\stage \brissot collects himself - revisiting a private mantra within his mind.

\startbrissot
Thank you, Monsieur \president.

\(beat)
Gentlemen, patriots are allowed to differ in opinions!

In fact, I say that this is healthy.

Because it is this same conviction we have for ourselves that we must
demonstrate against our enemies!

\(beat)
We should be very proud of how far we have come, brothers.

But we must examine our behavior and confess how disorderly and bitter we have
become.
\stop

\stage Applause.

\startbrissot[continued]
We fight among ourselves because we do not believe our fight for the revolution
is over.

We have not yet agreed on who it is we are supposed to be fighting, so we pick
from whoever is in the room with us.

\(beat)
Yet, despite this, the inevitable crisis we face at our borders goes ignored.

This is an existential threat to the fate of our nation, and therefore it is a
threat to the fate of liberty itself.

Fate demands that we take our fight to the Austrians - to crush the pride of
German princes!
\stop

\stage Applause. \condorcet gasps!

\startbrissot[continued]
A people who have won their freedom, after twelve centuries of slavery, must
have war to consolidate it!
\stop

\stage The \jacobins begin deliberating. \condorcet is heartbroken!

\startpresident
Silence! Decorum!
\stop

\startbrissot
Thank you, Monsieur \president.

\(beat)
Brothers, look to the Americans!

The Americans have braved the crucible of war and have continued the fight
since the moment they raised the banner of liberty!

Bless heaven that we are only two years late to the call.

Bless the time that it has given us in completing our constitution.

Now, it is a question of defending it.

A question of putting an end to those abroad who seek to destroy it.
\stop

\stage Applause.

\startbrissot[continued]
This shall be a war unlike any war in history!

For this shall be a war of LIBERATION!

Our soldiers will arrive as apostles of liberty; recruiting the oppressed to
the cause of freedom; fighting to the death any who interfere with the progress
of enlightenment!
\stop

\stage Enthusiastic applause.

\startbrissot[continued]
Now, let it be said, I welcome any patriot to defend their perspective to the
contrary.

But I implore them to examine my arguments and provide considered, healthy
refutations.

If I am mistaken, if I am convinced, I shall shock you all when I reveal my
embrace for their position.

\(beat)
But if you instead have your prejudiced misgivings, then speak them now.

For I swear on my life that I shall stand here to destroy your objections until
I drop dead.
\stop

\stage Pause. Whispers.

\startbrissot[continued]
I thought not.
\stop

\stage \brissot[thepatriot] dismounts the tribune to great applause.

\stage \robespierre closes his notebook and stands.

\startrobespierre
Monsieur! The floor!
\stop

\stage The applause continues.

\startrobespierre[continued]
Monsieur \president! I request the floor!
\stop

\stage The \jacobins cut their applause.

\startpresident
The floor is yours, \robespierre.
\stop

\stage \robespierre[theincorruptible] takes to the tribune.

\stage Pause. Silence.

\startrobespierre
Of any two opinions that split the \jacobins, one has on its side all the ideas
that flatter the imagination, all the brilliant hopes that ignite the spirit
and animate enthusiasm;

the other... is supported only by cold reason and difficult truths.

\(beat)
If I wished to please, I would defend the first.

If instead I wish to survive, I must support the second.

I understand if it shall displease you, but it is for the latter that I declare
myself.
\stop

\stage Pause. Silence.

\startrobespierre[continued]
Shall we make war, or shall we make peace?

Shall we attack our enemies, or shall we wait for them in our homes?

I believe that these questions fail to appreciate the depths of our conflict's
true character.

\(beat)
If the brilliant and prophetic painting of success truly ended in perfect
harmony and fraternal love for all the people of Europe,

and if this ambition is deemed worthy of something as extreme as declaring war,

then I would agree that Monsieur \brissot has perfectly resolved this debate.

\(beat)
Certainly, I would like as much as anyone to expand the reign of liberty - in
fact, I would gladly indulge in the pleasure of dreaming up its wonders with
you over dinner.

And if I were the master of France's destiny, I would have long ago sent armies
to neighboring tyrants to break the chains of their subjects.

\(beat)
But in the France I see now, I cast an anxious glare.

I ask: who is asking for war? How are they asking for it, in what
circumstances, and why?

\(beat)
I shall remind you, citizens, that all that matters is that we enlighten the
nation of its true interests and reveal to it the interests of its enemies.

So you ask: where are these enemies of the revolution?

They are among yourselves.

\(beat)
The bitter truth about this debate on war is that those who champion it are
doing so to advance their hidden agendas.

A scheme conspired by the internal enemies of liberty long ago.

This, citizens, is a trap.

\(beat)
Consider: who benefits from its outcomes?

Should France win the war, the court, having sanctioned it and who would have
executed it, would be treated as heroes.

This spirit and prestige would then, as it had for Caesar, be used to nullify,
then silence, the general will.

Should France lose, then we will be killed and, of course, the monarchy will be
restored.

\(beat)
So, I ask again... who is advocating for war? And why?
\stop

\stage Pause. Silence.

\startrobespierre[continued]
And, for what it's worth, nobody likes an armed missionary.
\stop

\stage \robespierre[theincorruptible] returns to his seat and opens his
notebook.

\startsplit

\startjacobins
Long live the nation!
\stop

\split
\leavevmode

\startjacobin[number=3]
Long live the incorruptible!
\stop

\stopsplit

\stage Great applause from the \jacobins\ - excluding \brissot.

\startcondorcet
Monsieur \president, I would like to request the floor.
\stop

\stage The applause continues.

\startcondorcet[continued]
I would like to request the floor.
\stop

\stage The applause continues.

\stage \brissot snaps out of a stupor.

\startbrissot
MONSIEUR!
\stop

\startpresident
Huh?

\(ringing a bell)
SILENCE! Decorum!
\stop

\stage \brissot gestures to \condorcet.

\startcondorcet
I would like to request the floor, Monsieur \president!
\stop

\startpresident
Oh! Yes, uh, the geometer wishes to speak!

\condorcet, the floor is yours.
\stop

\stage \condorcet[thephilosopher] takes to the tribune.

\stage Pause. Whispers. A few chuckles.

\startcondorcet
T-the citizens of - of France, having demonstrated their courage in claiming
their freedom, and having now demonstrated their calm firmness in - in
contemplating the danger that has just threatened it, no longer require
eloquence in the call to liberty.

It is therefore to their reason alone that we must speak on the security of a
peaceful future that is worthy of an enlightened people.

\(beat)
In the face of recent events, the question that we are all obliged to ask
ourselves is this:

is a king necessary for the establishment of liberty and the protection from
tyranny?

\(beat)
No, my friends, it is not.

\(beat)
The friends of royalty would argue that the power of an executive that is
established and limited by written law is far less formidable than that of a
usurped power that has no other limits than those of skill and audacity.

However, these people neglect to consider that a republic formed from a sound
and enlightened constitution, would see the powers of the army, the finances,
and the courts be adequately divided and distributed between disparate
departments and deputies.

Therefore, it would be necessary to adequately divided and distribut(ed) --
oh, I read this part already.

\(clears throat)
Uhm... t-therefore, it would be necessary to have destroyed, corrupted, or
denatured all these powers simultaneously before one could aspire to tyranny.

And since any enlightened constitution would guarantee the freedom of the
press, this threat is therefore impossible for a republic where deputies are
subject to frequent elections.

\(beat)
Consider that any popular tyrant can only act under a mask that hides their
hypocrisy.

Thus, the power ensured by the press to force these ambitious people to walk
with their faces uncovered shall render them all h-harmless.

Let us therefore not allow ourselves to harm the progress of the human species
and the happiness of our nation out of fear for an imaginary danger.
\stop

\stage Pause. Whispers.

\startcondorcet[continued]
No longer do we live in a time when laws can only be upheld through divine
intervention.

No longer are we prejudiced by the impious superstitions that suppose that any
man can speak on behalf of a supreme intelligence so long as he wears the
proper attire.

\(beat)
Yet, those friends of royalty, those who remember the events of history, but
who do not know history, remain frightened by the corruption of ancient
republics.

But let them examine these republics, and they will always see a division
between those who are sovereigns and those who are subjects.

Then they will understand the true forces that corrupt the privileged and
seduce the impoverished, and it will be clear that these forces cannot exist in
a nation where equality is made universal to all people who inhabit it - no
matter their race, their wealth, or their gender.
\stop

\stage The \jacobins are confused by this.

\startcondorcet[continued]
It is therefore not the tyranny of an individual that we must fear, but the
most odious and sinister tyrannies that exist when one multitude rules over
another through abuses of law and finance.

\(beat)
But is such abuse what the friends of liberty demand? Those who only accept
reason as the sole master of their fate?

I ask you: at whose expense could we satisfy the greed of our leaders?

What conquered provinces could a French general strip to buy our votes?

Would such an ambitious man propose to us, as he had to the Athenians, to levy
tributes on our neighbors in order to fund new temples and festivals?

Will he promise our soldiers, as he had to the Romans, the pillage of Spain or
Syria?

No, doubtless!

\(beat)
I ask you: how could we feed the greed of the few in a nation that refuses to
exploit their allies?

In a nation that refuses the conquests of war?

In a nation that refuses to be subjugated and built by citizens who refuse to
subjugate others?

\(beat)
I answer you, my friends, that it is because we refuse to be a kingly people
that we will remain a free people.
\stop

\stage \condorcet folds up his speech.

\stage Gradually, trepid applause builds into something polite.

\stopcomponent
