\startcomponent on_the_petition

\scene

\open Jacobin club. The \jacobins wait in silent anticipation. \robespierre is
casting his anxious glare.

\stage Enter \danton like a bull. The \jacobins applaud his entrance.

\stage Enter \brissot.

% TODO: Give \danton a speech.
\startdanton
Alright, listen up!

\(claps his hands)
It's about time to get to business.

The king is a traitor, we need him gone.

We at the Cordeliers wrote up a petition to request the king be put on trial.

\(tosses the documents into the crowd)
Read it on your own time.

Let me just hear your nagging.
\stop

\stage The \jacobins erupt into deliberation.

\stage \brissot is nervous. \robespierre shifts forward in his seat -
calculating.

\startbrissot
Long live \danton!
\stop

\startjacobins[cue={A Majority}]
Long live \danton!
\stop

\stage A great applause - excluding \robespierre.

\stage \robespierre stands.

\startrobespierre
Monsieur! The floor!
\stop

\stage The applause continues.

\startrobespierre[cue={The Incorruptible}]
I DEMAND THE FLOOR!
\stop

\stage Pause. Silence.

\startpresident
Oui, Monsieur \robespierre, the floor is yours.
\stop

\stage \robespierre paces about the club.

\startrobespierre
Citizens, we are allowing abstractions and fantasies to cloud our vision.

\(beat)
A petition? What's this talk of a petition?
\stop

\stage Pause. Silence.

\startrobespierre
What does it say? Who will sign it?

How many? Thousands?

Gathered where? At the Bastille? Or the Champ-de-Mars?
\stop

\stage Pause. Silence.

\startrobespierre
What interests do our enemies have?

They wish to put an end to conflict and an end to change.

And where are these enemies hiding?

Among the majority of moderates at the assembly, at the Paris municipality, and
in the National Guard.

\(beat)
How, I ask, could a peaceful petition of thousands of patriots gathered in one
location be a possible advantage for royalists with guns?

\(beat)
Who asks for this? And why?
\stop

\startdanton
I ask for it.
\stop

\stage The \jacobins gasp.

\stage Pause. Suspense.

\startdanton
What? DARE I NOT TAKE CREDIT FOR GLORY?
\stop

\stage Laughter. The \jacobins applaud.

\startjacobins[cue={A Majority}]
Long live \danton! Long live the revolution!
\stop

\stage Exit \robespierre[theincorruptible].

\stopcomponent
