\startcomponent prologue

\scene

\open The muffled chanting of La Prise de La Bastille can be heard from some
other room.

\stage Enter \condorcet holding his infant daughter.

\startcondorcet
What revelry! The sounds may be alien to your ears, but they signal the kind of
change that only the rarest of lifetimes can experience in the progress of
human history.

\(beat)
Oh, \eliza, how fortunate you are to have been brought into this world at the
perfect moment.

In this happy epoch, you are witness to the final embers of ignorance and error
as reason has emerged the victor in a battle so arduously fought by our
predecessors.

At last, we may cheer: "truth has prevailed; humanity is saved!"

And you - I am so proud to say - shall be the first to feel this great
benevolence!

\(beat)
To think that you have arrived just in time to be part of the first generation
of women to be educated in the same manner as men - to be the first of your
gender to assist the advancement of science as an equal...

There is no joy that compares to such a picture.

\(chuckles)
I admit, I dream too early.

There is work yet to be done, but... in this moment, just you and I, let us
bask in our shared capacity for hope.
\stop

\stage A knock at the chamber door.

\stage Enter \sophie.

\startsophie
\condorcet[personal], \brissot has arrived. And he brought a friend from
America!
\stop

\startcondorcet
Oh, he has?
\stop

\startsophie
Mhm!
\stop

\startcondorcet
Exciting! I'll be down in a moment.
\stop

\stage She prepares to leave but pauses to take in the sight. She giggles to
herself.

\stage Exit \sophie.

% CONSIDER: continuing \condorcet's dialog.

\stopcomponent
