\startcomponent recess

\scene

\open The office of \danton.

\stage Enter \brissot and the other \brissotins, minus \vergniaud.

\startbrissot
\guadet, check those drawers. \isnard, the desk. \gensonne, help \guadet with
the drawers.

\danton must keep records and evidence of his dealings here.
\stop

\stage The \brissotins each take their stations, they begin tearing up the
office.

\startgensonne
What are we searching for exactly?
\stop

\startbrissot
Anything that can give us leverage against \danton.
\stop

\startgensonne
Leverage for what?
\stop

\startbrissot
We need to know what \robespierre is planning.
\stop

\startguadet
What? Then why aren't we searching through \robespierre's office?
\stop

\startbrissot
Because he doesn't have one. He keeps all his notes in that little book he
carries with him.

\beat

Now, shut up and search. Receipts, letters, contracts, notes, anything!
\stop

\startisnard
Wait, where's \vergniaud?
\stop

\startbrissot
What concern is that to you?
\stop

\startguadet
He was just asking, \brissot.
\stop

\startbrissot
You mean, instead of searching?

\beat

\vergniaud? He's still working on that speech of his.

He's na\"ive. The crowd cannot be won from \robespierre through something as
immaterial as ideology.

We need proof!

Proof of their treachery!

Proof that \danton and \robespierre are plotting to seize control of Paris!
\stop

\startgensonne
Ridiculous!
\stop

\startbrissot
Ridiculous?
\stop

\startgensonne
What makes you think they aren't loyal to the people as they claim?
\stop

\startbrissot
Of course he says he's loyal to the people! What politician wouldn't say such
a thing!

I know \robespierre, he's a damn serpent.

What matters is who he is when he is alone with \danton and that... dreadful
\marat...

And can you really look me in the eyes and tell me that he isn't plotting to
make himself the new king of France?
\stop

\startgensonne
And \danton? He's beloved by the citizens~- more so than \robespierre.
\stop

\startbrissot
By god! Enough questions! You sound just like \condorcet...
\stop

\startguadet
I think I found something.
\stop

\startbrissot
What is it?
\stop

\stage \brissot takes the document.

\startguadet
It's reimbursement for courier fee's and sealing wax for documents that had
been delivered to the lodgings of \robespierre.
\stop

\startbrissot
I can see that.

This won't do. It tells us nothing we don't already know!

Of course the two are in conspiracy together!
\stop

\startsplit

\startguadet
Yes, but~--
\stop

\split
\leavevmode

\startbrissot
-- Move aside.
\stop

\stopsplit

\stage \brissot begins searching at \guadet's station.

\stage \isnard and \gensonne stop their search and watch \brissot.

\startbrissot
Ah! Perfect!

A letter without a signature, but addressed to the simple "citizen."

This bears all the signs of \robespierre.
\stop

\startisnard
But you say the letter is unsigned?
\stop

\startbrissot
Who else do you know that uses the word "citizen" to address patriots?
\stop

\startgensonne
I'd say plenty.
\stop

\startbrissot
But who among them would also wish for a veil of anonymity in correspondence
with \danton?
\stop

\startguadet
What does the letter discuss?
\stop

\startbrissot
"Citizen, it is my wish to bear thanks to you for the integrity of your words,
the boldness of your actions, and the strength of your success."

\beat

Clearly this is evidence that \robespierre was behind the insurrection of the
Tuileries!
\stop

\startgensonne
The letter is unsigned!
\stop

\startbrissot
Who else could have written such a note?
\stop

\startgensonne
Any one of the citizens who celebrated the defeat of the Tuileries!
\stop

\stage Enter \vergniaud.

\startvergniaud
\brissot! What in heaven are you doing!?
\stop

\startbrissot
Get out of here, \vergniaud. You've made your choice to not contribute to the
defense of liberty.
\stop

\startvergniaud
You call the ransacking of the Minister of Justice a justice for your country?
\stop

\startbrissot
You've seen what barbarity monsters like \danton are capable of!

You may not like it, \vergniaud, but this is a {\emph necessity}!
\stop

\startvergniaud
Ha! "Necessity" is the excuse of tyrants. You debase yourself.

\beat

All of you! All of you debase yourselves by following in this example.

For shame!

Does the voice of true justice not burn within you at the sight of this
degeneracy?
\stop

\startbrissot
Don't talk to them like you know better, \vergniaud.
\stop

\startvergniaud
My apologies, \brissot, I wasn't aware I was encroaching on your influence.

\beat

\guadet! Do you concur? Have I interrupted the march of liberty in being here?
\stop

\stage Pause.

\startvergniaud
\isnard? And you?
\stop

\stage Pause.

\startvergniaud
And you, \gensonne...?
\stop

\stage Pause.

\stage \vergniaud shakes his head and sighs.

\startvergniaud
If only I could have learned the magic spell that has been cast over your
hearts...
\stop

\stage Exit \vergniaud.

\startbrissot
What insolence. What ignorance.
\stop

\startisnard
Insolence? \vergniaud is the best of us and you know it.
\stop

\startbrissot
There are no best and worse when we are all patriots...

And you know it.
\stop

\startdanton[offstage]
WHAT!? THAT FILTHY FUCKING RAT!
\stop

\stage Enter \danton~- followed by \vergniaud.

\startbrissot
\vergniaud, you self-righteous bastard!
\stop

\startdanton
YOU'RE A DEAD MAN!
\stop

\stage \isnard, \guadet, and \gensonne retreat behind \vergniaud.

\stage \danton pulls out his pistol! \brissot dives behind the desk as \danton
fires!

\stage \danton tosses his gun aside and charges at \brissot.

\startbrissot
\beat{during chase}

You know he's a monster!
\stop

\startdanton
I have use of monsters.
\stop

\startbrissot
What is he planning!?
\stop

\startdanton
You think he tells me shit?
\stop

\stage \danton punches \brissot, causing him to collapse to the ground.

\startvergniaud
I'm sorry, \brissot. But~--
\stop

\startbrissot
\beat{to \danton}

-- WHAT IS HE PLANNING!? TELL ME WHAT HE'S PLANN(ING)~--
\stop

\stage \danton punches \brissot another time, then lifts him up by the collar.

\stage Pause.

\startdanton
\robespierre was right about you...

\beat

That damn mask you wear...
\stop

\stage \danton throws \brissot over to the \brissotins.

\startdanton
Get the fuck out of my office.

All of you!
\stop

\stage Exit all but \danton.

\stage Pause.

\stage \danton walks over to a picture that hangs on the wall, opening it to
reveal a hidden safe.

\stage To his relief, the contents have not been touched.

\stage Exit \danton.

\stopcomponent
