\startcomponent tale_of_a_revolutionary

\scene

\open The \vernet residence living room.

\stage Enter \vernet.

\stage She sets the table with a breakfast consisting of coffee and an
omelette.

\stage Enter \eliza.

\startvernet
Good evening, dear! I've prepared you breakfast!
\stop

\starteliza
It's near daybreak.
\stop

\startvernet
You're mistaken, it's near sunset.
\stop

\startvernet
Dear, you look like you are still living in yesterday.

I can see it in your sunken eyes.

\beat

Oh! What am I saying.

\beat{embraces \eliza}

I shouldn't make such comments about a woman's appearance.

I apologize, darling. I made you breakfast.
\stop

\stage \eliza seats herself.

\starteliza
This omelette is enormous. How many eggs did you put in it?
\stop

\stage \vernet seats herself.

\startvernet
Twelve.
\stop

\starteliza
What? Twelve?

\beat{chuckles}

My attention must be worth a fortune.
\stop

\startvernet
It is to me, dear. Even if through such tired eyes.
\stop

\stage \eliza begins her meal by taking a sip of coffee.

\startvernet
I wish to apologize.
\stop

\starteliza
Again? I think it's an astute observation.
\stop

\startvernet
For my mood at where we had left off in my tale.
\stop

\starteliza
Oh.
\stop

\startvernet
I have reflected on this matter and have decided that its cause was a selfish
desire to acquire knowledge that is forbidden to me.

Please understand, dear: I am an old woman.

Many of my days are occupied in my fantasy of completing the stories most dear
to my heart.

Thus, I reacted with hostility when your story failed to live up to these
expectations.

\beat

I apologize, dear.
\stop

\stage \eliza smiles.

\stage \vernet smiles. She begins stuffing a pipe with tobacco.

\starteliza
Please, madame. I don't mind displays of passion when they are honest.

In fact, I hope to experience whatever passions remain from your story~-
incomplete as it may be.
\stop

\startvernet
\beat{while lighting her pipe}

Oh, yes, dear. You have a debt to me.
\stop

\stage \vernet chuckles, puffing up the room with smoke signals.

\startvernet
When your mother first introduced herself to me, and informed me of her
objective, I was convinced that the only reasonable cure your mother could of
had in mind was a night of passion.

\beat{puff}

Thus~- O, woe! I become swept in a frivolous passion for respecting the privacy
of these lovers!

At once, I compel \marcoz, \sarret, and myself to banish ourselves until
sunrise.

What foolishness...
\stop

\starteliza
So you asked for details once~--?
\stop

\startvernet
-- Don't interrupt, darling.

I am an oracle~- I speak on behalf of the Parcae as they twine the thread of
life.
\stop

\starteliza
\beat{chuckles}

As you say, madame.
\stop

\startvernet
When we return in the morning, the three of us are shocked to see this living
room covered in pamphlets!

Your father, sprawled along his doodles like a schoolboy, explains that your
mother had left prior to dawn to satisfy an alibi come daybreak.

He then proudly proclaims to have discovered the existence of what he calls a
universal calculus;

insisting that such a calculus is powerful enough to represent any thought
reachable through reason, and that it was necessary that this discovery be
preserved for the pursuit of human progress.

\beat{puff}

I am astonished.

\beat{puff}

"Dear, lord," I think...

"When the woman meant reason, she meant {\emph arithmetic}!"

\beat

He then assures us that the discovery itself is secondary to the importance of
introducing such novel ideas to the public.

He says your mother explained that how an idea is presented is exactly as
meaningful as what is presented.

So, he announces that by spring he shall have written his great tableau of
human history.

The magnum opus for which he plans to prove that the flow of human progress is
not random, but is instead driven by the indefinite perfectibility of reason,
science, and education.

\beat

Doubtless, you have read some of this tableau last night.
\stop

\starteliza
Actually, I have been reading letters first.
\stop

\startvernet
Darling, it takes effort to pluck a page at random and find anything not
belonging to his tableau.
\stop

\starteliza
What are you accusing me of?
\stop

\startvernet
Did you not so much as glance at how the ink sings and flows on the page?

You sorted them, yet failed to parse even a single phrase long enough to notice
the joy of its prose?
\stop

\starteliza
I sorted them by page number.
\stop

\stage \vernet snobbishly pouts, then stands to fetch more tobacco from a
nearby tin.

\starteliza
You scoff? I fail to see the crime in reading chronologically.
\stop

\startvernet
Oh, I accuse you of nothing, dear.

I simply don't believe you.
\stop

\stage \vernet reignites her pipe~- another round of puffs as she returns to
her seat.

\startvernet
In fact, I say it's adorable that you mention this.

You see, your mother wasn't the only visitor your father had during his time
writing his tableau.

\beat{puff}

Your father had the pleasure of teaching you to count!

O, you were such a little angel!

And, ah! I remember that stunned face you had when you first recognized your
father through that terrible beard he had grown!
\stop

\starteliza
Yes, my mother told me of this.
\stop

\startvernet
But you don't remember?
\stop

\starteliza
I remember how to count.
\stop

\stage \vernet considers this solemnly.

\starteliza
And have retained a long-felt repulsion for beards.
\stop

\startvernet
\beat{puff}

Mm. Good girl.

\beat{puff}

You came to visit me just prior to the harsh winter of seventeen ninety-four.

Such a terrible season...

Many died to lack of food and firewood.

\beat

Your mother writes during this time that she sustains the family by trading for
beans and lentils.

\cabanis, on occasion, comes to visit with prescriptions for your father's
worsening leg~- injured, no doubt, by all that sitting.

\beat{stands}

Speaking of: would you like some more coffee, dear?
\stop

\starteliza
I'm not yet finished with my current cup.
\stop

\startvernet
Then certainly it has gotten cold by now.

I shall return with the remedy.
\stop

\stage Exit \vernet.

\stage \eliza stands and surveys the room once again.

\stage She studies the entrance, her eyes trace the path a child once walked.

\stage She squats to align herself with that child.

\starteliza
\beat{softly}

I don't remember...
\stop

\stage Enter \vernet with two cups of coffee, causing \eliza to snap back into
her adult stature.

\startvernet
What are you doing, darling?
\stop

\starteliza
I developed a fear of sitting while you were away.
\stop

\stage \vernet laughs.

\startvernet
You're far too young to worry about such things.

Come sit. Fresh coffee.
\stop

\stage \eliza returns to her seat and partakes in the coffee placed before her.
\vernet remains standing and strolls about the room.

\starteliza
You are quite fond of coffee.
\stop

\startvernet
\voltaire recommends at least twelve cups a day.
\stop

\stage \eliza chuckles.

\starteliza
Probably not the most trustworthy recommendation.
\stop

\startvernet
I've had no trouble with it.
\stop

\starteliza
You were last discussing the winter of seventeen~--
\stop

\startvernet
-- Ah, yes. The winter of seventeen ninety-four.

Terrible season that was.
\stop

\starteliza
You've mentioned that.
\stop

\startvernet
Oh, of course I had. Many died that winter to~--
\stop

\starteliza
-- To hunger and cold.
\stop

\startvernet
Mm, enough of that, dear.
\stop

\starteliza
Apologies. I had forgotten that the fates are speaking.
\stop

\startvernet
That's right.

\beat

Hunger and cold claimed many, yet what was the solution offered by the
\convention?

Justice, they called it. The order of the day.

Speedy, severe, inflexible justice.

\beat

Thousands of live bodies were hauled, cart by cart, to the Place de la
R\'evolution, only to then be carted to the edge of Paris to be forgotten in
unmarked graves.

Speedy, severe, inflexible.

That abominable machine.

\beat{deep breath}

I never asked for anything from your father in exchange for the protection I
offered him from this machine.

But there was something invaluable he provided me in those cold winter nights.

You see, each night, when your father had concluded his writing for the day, he
would emerge and sit exactly where you are now.

There he offered \sarret and I stories of history that would always bring us
peace.

All told with a style that was unique to the night's chosen tale, all poured
from an infallible well of memory only a lifelong scholar could possess.

\beat{chuckles}

While men toiled in riotous debates about virtue and terror, it was a great
relief to listen to this scholar rant about the controversies of ancient
science and the cruelties of ancient wars.

One of my favorite tales was of a potter~- Palissy I believe the name to be~-
who dedicated sixteen years of his life in his attempt to reproduce a single
porcelain cup originating from China.

Going as far as reducing his family to poverty, denying his children their
suppers, and even burning their furniture to keep his furnace lit.

\beat

He never successfully reproduced the cup.
\stop

\stage \eliza smiles as she scoffs.

\stage \vernet begins relighting her pipe.

\startvernet
\beat{puff}

Then, as winter neared its end, so too did his tableau.

I had begun to notice that the stories he were telling were arranged
chronologically throughout history.

Thus, when he began telling stories near the present, I was obviously worried.

Yet, to my delight, his stories were not of violence and virtue, but rather
about \voltaire and his eloquence.

About his fascination with \leibniz and the nature of language and knowledge.

About his mentors, Turgot and d'Alembert.

And the influence of Newton on the principles of physical laws...

\beat{puff}

And... when he finally approached today, he began his story differently.

With an apology, of all things.

I know not why.

\beat{puff}

He says that he believes that all human beings are endowed with a mind
capable of an indefinite perfectibility;

that which we have observed by recounting the progress of our history;

and by which we may sketch, without fear and with some pretense of truth, the
future destiny of humanity on the basis of its history.

\beat{puff}

"This is going to be good one," I say.
\stop

\stage \eliza chuckles.

\startvernet
He then tells me this tale, and I shan't ever forget it:

\beat{puff, sets down the pipe}

In the city of New Atlantis, infinite productivity and peace had been achieved
by the discovery that all problems that are difficult, are difficult for the
same reason.

As a consequence of this discovery, citizens no longer believed anything to be
real until it could be adequately communicated using a shared interpretation of
symbols.

And in New Atlantis, these symbols are used to inscribe laws onto one of three
colors of parchment: red, green, and blue.

\beat

Physical laws are colored blue, and these describe the laws that all citizens
obey by not acting.

An example of such a law would be: "the choice to imagine peace is the act of
creating it."

\beat

Green parchment is reserved for "safe" laws.

These were laws that have been proven by evidence to reduce the difficulty of
labor.

An example of such a law would be: "all people have seven holes for seeing,
hearing, eating, and breathing."

\beat

And lastly, and the most infamous, are red colored laws.

These are "unsafe" laws: or, laws whose interpretation is contingent on a
secret understanding known only by two people.

An example of such a law would be:

\beat{eye contact with \eliza}
"I shall gladly, blindly, patiently obey all of New Atlantis' known laws."
\stop

\stage Pause.

\starteliza
And that's the end of the story?
\stop

\startvernet
Yes. The ending just doesn't make sense to you yet, darling.
\stop

\starteliza
Is that not the point of a story? To understand the ending?
\stop

\startvernet
That is the point of most stories.
\stop

\starteliza
But not all of them?
\stop

\startvernet
Not this one. Not yet.
\stop

\starteliza
Which one?
\stop

\stage \vernet smiles.

\startvernet
Enough interruptions, dear.
\stop

\starteliza
Sorry.
\stop

\stage \vernet nods. Then picks up her pipe to refill it with tobacco.

\startvernet
Once spring had arrived, your father had completed his tableau with time to
spare to make additional corrections to his universal calculus.

He also made time to provide tutoring for \sarret in new theories on the
subject of integral calculus.

\beat{sits}

However, as their tutoring became more frequent, they eventually began speaking
to one another in Latin.

\beat{ignites the pipe, puff, puff}

I begin to suspect conspiracy.

\beat

I ask myself: what does a scholar, who can write for two seasons on end, do
when he completes his great work?

He claims to have already invented the universal language of science, so what
else is there for him to invent next?

What will remain to satisfy him if he were to remain entrapped within my home?

\beat{puff}

Perhaps, he will wish to make himself a martyr?

\beat

Then letters from your mother begin to arrive that bring terrible news: she
says she is being visited by inquisitive men asking about the whereabouts of
her husband.

She fears for her safety and she proposes divorce.

Ah! I know!

\beat{puff}

Oh, and of course, your father is devastated.

And worse yet, it reignited his fear of death~- this time, not of his own~- but
of your mother's... and of yours...

\beat

But to good fortune, he came quickly to his senses, and demonstrated a new
courage founded in the instincts he had for his family's security.

He agreed to divorce, yet this wasn't necessary: he had no need to be present
for the proceedings, or provide his consent.

In fact, this was made possible due to the laws inherited from the Legislative
Assembly.

Laws he had written himself.

\beat{puff}

Soon, however, strangers would begin visiting my door.

\beat

Usually these visits would come under the pretext of seeing about a vacant room
to rent~- not unlike the visit I received from \cabanis that fateful night.

However, these men always spoke with a certain rhythm that I did not hear in
your uncle.

Instead, in their voices, I hear a cadence that revealed a certain...

\beat{as if just discovered}

They spoke as if they were actors in a play.

\beat

And your father, no doubt, overhears these strangers and doubtless hears this
cadence, too.

I was sure that his instincts would again be ignited.

This time, for the safety of myself.

\beat{puff}

His second mother.

\beat

Thus, when I hear Latin between himself and \sarret, I become increasingly
paranoid of conspiracy.

\beat{puff}

For sure, he plans to escape my prison of life.
\stop

\stage Pause. \vernet sets down the pipe.

\starteliza
But why would you suspect that? That would be foolish.
\stop

\stage Pause.

\startvernet
I suspected that it was his instinct to die on a battlefield: that it was
always better for him to die alone, than for him to die with his family.
\stop

\starteliza
Did he succeed?
\stop

\startvernet
Yes. Eventually.
\stop

\starteliza
And he was guillotined.
\stop

\startvernet
No.
\stop

\starteliza
No?
\stop

\startvernet
Dear, despite his fame, your father was still a philosopher recognized only by
his name.

Hardly anyone in Paris could recognize him by sight.

He was a stranger to them.
\stop

\starteliza
So he simply wandered Paris? Homeless?
\stop

\startvernet
With that beard and bad leg. Yes. He certainly played the part.
\stop

\starteliza
Where did he go?
\stop

\startvernet
He ended up outside the city, in the countryside to the southwest.

I suspected that he spent the nights sleeping under the stars.
\stop

\starteliza
What did he do for food?
\stop

\startvernet
Mm. Tragic that he must have it.

This was how he was caught by the local surveillance committee.

An innkeeper became suspicious of a homeless man when he suddenly ordered a
twelve-egg omelette.

He was asked to prove his identity, and when he failed, he was arrested.

\beat

But nobody recognized him. Nobody recognized the philosopher.
\stop

\starteliza
So what did they do with him?
\stop

\startvernet
He was imprisoned.

And he was found deceased in his cell the following morning.
\stop

\starteliza
What? No. How?
\stop

\startvernet
I suspect that it was during your uncle's visit with him before the start of
winter.

I believed he provided your father with hemlock, and~--
\stop

\starteliza
-- What? No. That's absurd.
\stop

\startvernet
Excuse me?
\stop

\stage Pause.

\starteliza
He killed himself?
\stop

\stage \vernet gestures to respond, but chooses silence instead.

\starteliza
Why would he do that?
\stop

\stage Pause.

\starteliza
No. That. I don't understand.
\stop

\startvernet
What don't you understand, darling.
\stop

\starteliza
Why would he choose that?

That.
\stop

\startvernet
I'm very sorry, dear.
\stop

\starteliza
No. He wouldn't have.

Why?
\stop

\stage Pause.

\stage \vernet embraces \eliza.

\starteliza
WHY? Why? So he~-- so he could pretend to be Socrates!?
\stop

\stage Pause.

\startvernet
It's alright. Shh...
\stop

\stage \eliza takes a deep breath. Gone.

\starteliza
My apologies.
\stop

\startvernet
Oh. Never mind that.

I don't mind displays of passion when they are honest.
\stop

\stage Pause.

\startvernet
Let me fetch you some coffee.
\stop

\starteliza
No, thank you. I have had enough.
\stop

\stage Pause.

\starteliza
I also wish to speak to you about tenancy.

There are more documents than I expected and it will take~--
\stop

\startvernet
-- Darling... Have you learned nothing from my tale?

\beat

You can stay for as long you need.
\stop

\starteliza
I don't have money for rent.
\stop

\stage \vernet laughs.

\startvernet
Nonsense. I only ask one thing of you.
\stop

\starteliza
What would that be?
\stop

\startvernet
Repeat after me:

\beat{eye contact}

"I shall gladly, blindly, patiently obey all of \vernet[madame]'s known laws."
\stop

\stage \eliza chuckles.

\stage \vernet is serious.

\starteliza
"I shall~--"
\stop

\startvernet
-- The eye contact is important, dear.
\stop

\starteliza
I shall gladly, blindly, patiently obey all of \vernet[madame]'s known laws.
\stop

\startvernet
Mm. Good girl.

\beat

Now, allow me to fetch myself some coffee!
\stop

\stage Exit \vernet.

\stage Pause.

\stopcomponent
