\startcomponent the_mountain

\scene

\open The Salle du Man\`ege. The \convention[national] has reconvened. A
\deputy, by the name of \salles, is positioned at the tribune.

\startsalles
Six hours ago, you have acted as jurors in pronouncing the guilt of
\louis[xvi].

And now... you invent laws as legislators so that you may pass sentence onto
him as judges?

By what logic has brought us to abandoning logic altogether?

By what standards do we have, that yet remain, to hold ourselves accountable to
our actions?

\beat

By my ear, this convention has observed that the fate of \louis presents... a
political question: that the matter of sentencing him is an imperative to the
interests of the state.

\beat

It's true: if \louis is to deserve death, and does not suffer it, he will
remain a constant source of discord.

Any assassin henceforth may announce on the scaffold: "why am I condemned to
this fate? Have I committed a crime worse than the betrayal of my country?"

\beat

But should \louis be executed, the pretensions of the royalists shall survive
him.

The death of a monarch does not signal the death of a monarchy.

And it is this embarrassing reality that manifests the dilemma between the two
opposing views of this assembly.

\beat

The regrets of the past, the evils of the present; a foreign invasion, public
unrest; and the dissension between ourselves.

All of these threaten us with the greatest misfortunes.

But are we to blindly continue the path traced before us by the factious men of
Paris?

\beat

Already, some bold speakers have established the disgusting doctrine that all
men are born with the inalienable right to assassinate \louis.

Already, these factious men have sanctioned the assassination of a thousand
prisoners.

And now... the \convention has inexorably been accused of sharing these
bloodthirsty maxims because the evil few within won't cease shouting it.

\beat

Well! What shall we think of these indiscretions? What character do they assume
in my sight?

What representatives continue to invoke the axe of the people against the
\convention itself?

Threatening death to those who fail to deliver death upon \louis?

MURDERERS! What shame!
\stop

\stage The \rightminority applaud.

\startbarere
Order!
\stop

\startsalles
O, my country!

\beat

I say there is only one method to avoid such great evils.

I say that the \convention, having pronounced the guilt of the regent, should
submit to the people the punishment he ought to suffer.
\stop

\stage \salles dismounts the tribune to great applause from the \rightminority,
and great disdain from the \leftminority.

\startserres
I request the floor, \barere[citizen].
\stop

\stage \barere affirms as --

\startmarat
-- NO! The damned moderates have had their turn!
\stop

\stage The \rightminority viciously protest.

\startrightminority
Murderer! Tyrant!
\stop

\startmarat
HOW DARE YOU CALL ME A TYRANT!
\stop

\startbarere
Silence! I WILL have you removed from this assembly if you continue these
damnable outbursts.
\stop

\startmarat
SEE!? SEE!? AND YOU CALL ME THE TYRANT!?
\stop

\stage \serres ascends to the tribune.

\startserres
I have no desire to justify the crimes of the accused.

I do not come to raise your compassion for his unhappy situation.

I (have) --
\stop

\startleftminority
-- Royalist!
\stop

\startserres
I have come to support what I believe to be necessary for the heart of
liberty and safety.

I only ask you: what is the greater danger to the state, the circumstance of
the king's death, or that of his imprisonment?

\beat

I will not insult the French by supposing that the restoration of the King will
eliminate hatred of royalty.

However, I do suppose that his existence, in my opinion, ought to create no
alarm.
\stop

\stage Vicious hatred from the \convention.

\startserres
BESIDES! BESIDES!
\stop

\startbarere
ORDER! SILENCE!

\beat

Come on. You all know the same rules apply to everyone.
\stop

\startserres
Besides... you all know that the people of France believe him to have a gentle
heart - that his faults are due to his frailty more-so than his will, and that
he was actuated by a woman who knew how to captivate him.

And after all, if \louis is executed, let us not forget his young heir.

Not only would his execution fail to remove the monarchy, but his children
shall be devastated!

And can any of you dare to argue that a child should be punished so severely
for the crimes of their parents?

\beat

I therefore insist that \louis be imprisoned until the conclusion of the war.
\stop

\stage Incredible jeers from the \convention as \serres returns to his seat.

\startguadet
This man is no moderate! He speaks only for himself!
\stop

\startgensonne
\guadet, shut up!
\stop

\startguadet
The decision should be settled by the nation!
\stop

\startmarat
This is clearly biased! Clear favoritism!

\beat{to \barere}

YOU, MONSIEUR, ARE A BIASED WHORESON!
\stop

\stage The \leftminority applaud \marat.

\startmarat
I REQUEST THE FLO-O-OR!
\stop

\startbarere
Denied for lack of decorum in addressing my person.
\stop

\startmarat
Lock them up! LOCK. THEM. U-U-UP!
\stop

\stage The \rightminority shout at \marat.

\stage The \leftminority look to \robespierre as if they expect something to
happen.

\stage He does not move.

\startvergniaud
I request the floor, \barere[citizen]!
\stop

\stage The \rightminority applaud!

\startbarere
\vergniaud, the floor is yours.
\stop

\stage The \leftminority jeer.

\startbarere
Enough!
\stop

\stage \vergniaud[theeagle] ascends to the tribune.

\startvergniaud
By what means do we define the sovereignty of the people?

This sovereignty gifted with the power to construct law?

Our enlightenment answers that it is the seeking of improvements to the
happiness of society that the people, sovereign in their own domains, render
sovereign the collective of their country.

These people exercise this power either themselves or through their
representatives, and it is in this latter case, the one for which we find
ourselves, where decisions of public welfare are made into law.

\beat

But how can this be so? How is it that a minority of representatives can speak
on behalf of the majority of their nation?

\beat

Because these representatives are presumed to be the expression of the general
will.

Their validity is manifest by this assumption alone - and by this assumption,
they are granted clemency for the power they wield.

\beat

Yet, by the uncertainty inherit in all assumptions, including the one just
observed, the people must always reserve the right to correct their
representation~-

to make manifest their true will when this presumed will is found to be
imprecise.

And I say that it is self-evident that any attempt to deprive the people of
this right is to deprive them of their sovereignty.

\beat

And now I ask you: when I speak of this people, who is it that you picture?

Is it the people of France that you imagine? Or rather, is it the people of
Paris?

\beat

The sovereignty of the nation is that of the whole - not merely that of its
largest metropolis.

And thus, to insist that some Parisian upstarts speak on the behalf of all of
France - to insist that those men who have seized the Tuileries on the tenth of
August have, in doing so, successfully seized the general will...

\beat

This belief, gentlemen, is nothing short of criminal.
\stop

\stage The \leftminority violently protest!

\startbarere
BY GOD, YOU SHALL STAY IN YOUR SEATS, CITIZENS! ORDER!
\stop

\startvergniaud
I promise you, gentlemen, that the general will shall be heard!

Indeed, you shall be heard!

You unhappy Parisians! I harbor a deep love for your incredible passion for our
country!

And I assure you that patience for opinions will not be our doom!

\beat

But these troubles, these misfortunes that I continue to hear about - and have
excited these extraordinary passions - have manifested exclusively within Paris
alone.

Ask yourselves, you amongst you who orchestrated the tenth of August, have any
of you even set foot outside of Paris since the beginning of the war?

Have you seen with your own eyes the barbarity you read exacerbated in the
journals and broadsheets?

I assure you... beyond the veil of these Parisian borders is a world that is
calm.

A world full of people who wish only for the stability of sound laws and
reasoned justice.

\beat

But is this what I hear being offered by the factious men of Paris?

By the assassins who execute arbitrary laws decided by whims and superstitions?

Those {\emph dictators} who seek to cannibalize my beloved country!
\stop

\stage Applause from the \rightminority.

\startvergniaud
Alas! Such tragedy that your ears have been dinned by the ideology of terror!

You have been manipulated by these decriers.

They have been fanning the flames of your most savage instincts so that the
ideologues who are embarrassed by their own ignorance do not need to feel
ashamed for having lost all sense of virtue.
\stop

\startsplit

\startbrissot
HEAR, HEAR!
\stop

\split
\leavevmode

\startgensonne
Hear, hear!
\stop

\stopsplit

\stage \robespierre sneers - never has a man hated another so deeply.

\startvergniaud
It is true that there are men whose every breath is an imposture, as it is the
nature of the serpent to exist only by the distillation of venom.

You have heard accusations laid against our country - that our people consist
of an evil within - of a majority composed of schemers, of aristocrats, and
so-called moderates.

But the bold-faced truth is that these accusations are laid against any who are
not sufficiently radical by the standards of snakes.

The ideology of terror has been weaponized against you - resorting to
radicalization as the only means it has for justifying the HIDEOUSNESS of
this sanguine insanity!
\stop

\stage The \rightminority cheer.

\startvergniaud
Unhappy Parisians! You who are so worthy of universal liberty, what will become
of you when Paris is deprived of the sane?

What industry can survive when the industry of slaughter becomes total?

What labors will remain to sustain your families?

Massacres do not grow wheat. They do not distribute grain. They do not bake
bread.

And yet! When the day comes that you must watch your family beg and starve,
what will these perfidious monsters offer you?

These false prophets who dare call themselves your friends?

They will point you to the quarries filled with corpses and exclaim:

"you are oppressed by hunger - dine, then, upon the fruits of my labor!"
\stop

\stage The \convention recoils in horror.

\startvergniaud
You shudder, citizens; O, my country!

I speak of such extremes, but let history record my effort to prevent this
terrible fate.

In this persuasion I remain steadfast in my observation of an irrefutable
truth:

that any act of a representative that attempts to subvert the formal or tacit
approval of the people is an enemy to their sovereignty.

The people alone, who a year ago has promised the inviolability of \louis,
reserve themselves the exclusive right to revoke this protection in the light
of his crimes.

If it is true that the people desire his death, then immediately so it shall
be.

If instead they are of the contrary and you take it upon yourselves to override
them, then I do not pity you for what terrifying burden shall be brought upon
your heads.
\stop

\stage Pause. The \leftminority hiss.

\startvergniaud
I have nothing more to say...
\stop

\stage \vergniaud dismounts the tribune to incredible applause from the
\rightmajority!

\startmarat
ENOUGH OF THIS!
\stop

\stage \marat[thefriend] seizes the tribune.

\stage He squares up to the \convention\ - turning up his chin to the ensuing
shouts and accusations.

\startrightminority
THE TYRANT! THE DICTATOR! THE TRAITOR!
\stop

\startbarere
Order! ORDER-R-R!

\beat

\marat, just get on with it.
\stop

\startmarat
Oh, is it my turn now to speak?

\beat{spits}

I shall not cower to your threats or bend to your biased proceedings!

You accuse me of wishing for a dictatorship?

No longer do I say this! No longer do I say to install a dictatorship!

I have told you this many times, yet leave it to traitors to deny me my right
to change my mind!

\beat

I listen to this violence and vitriol laid against my person, and I have become
convinced that FIFTY YEARS OF ANARCHY AWAIT US!

And after this, you shall see the very dictator you fear most emerge - a true
patriot and statesman will reveal himself to crush the weakness of greed and
corruption!

Yet what disgusting hate must I continue to suffer in order to convince you of
this?

I am no longer a believer in dictatorship! I wish to avoid this!

If only your hate would cease to slow me down!

This party of \girondins are a menace to humanity for denouncing my plans to
save our nation!
\stop

\startguadet
You're the party of murderers!
\stop

\startmarat
SHUT UP! SHUT. UP! HOW DARE YOU INTERRUPT ME!

I sat there like a FUCKING ANGEL listening to your precious hero accuse good
patriots of cannibalism!

Oh, but now that poor, foolish \marat is speaking, you get to play the part of
saints? Of soldiers for justice and liberty?

Hypocrites and liars - ALL OF YOU!

\beat

Good citizens and patriots, do not listen to these scheming aristocrats and
degenerate moderates!

They are deceivers!

You say that the France beyond Paris is calm? Tranquil? What horseshit! What
ignorance!

They are clearly still in denial that the war they fought so hard to plunge us
into is anything but the swift victory they had promised you!

Wishing in vain to paint a picture of a "polite invasion" of the French
provinces!

\beat{spits}

As if they deliberately wish to erase reality from the blood and ash that coat
Verdun!

To wash clean the wombs of raped women!

To wash away the ugly bits of skull and brain that once belonged to Colonel
Beaurepaire's bastard head!

\beat

Good citizens, loyal patriots, do not listen to their hate. Even their hate is
a lie!

They are nothing but talented actors who believe you too stupid to see through
their false passions!

Think about it!

Only a man of the most sinister and villainous quality could possibly shake
your convictions with a speech that he DIDN'T. EVEN. WRITE. HIMSELF!
\stop

\stage The entire \convention stands in shock!

\startleftminority
FRAUD! FRAUD! {\WORD\vergniaud} IS A FRAUDULENT ORATOR!
\stop

\startcondorcet
WHAT!? No! Reason is independent of the speaker!
\stop

\startmarat
HA! A CONFESSION! THE MASK OF PATRIOTISM SLIPS!
\stop

\stage A cacophony of insults are laid upon the \rightminority\ - a devastating
scandal!

\startmarat
I SAY! I say that no man can be trusted whose words do not come from his heart!

Whose words do not derive from the true passions of their own soul!
\stop

\stage Applause from the \leftmajority!

\startmarat
We must purge France of this disease of moderates! To save France from the
brink of anarchy!

I motion for the formation of a new committee!

A committee dedicated to the cause of public safety!

A committee comprised of only the best and most loyal patriots in the nation!

This committee will solve all the troubles facing France - a final solution for
liberty!

Our deliverance will be swift - it will inspire awe!

So swift it will be that this committee would only need to exist for a SINGLE.
DAY.

A SINGLE DAY I TELL YOU!

And by sunset, ALL of France's problems will be solved at once!
\stop

\stage Outcry from the \rightminority. Applause from the \leftmajority!

\startguadet
{\WORD\marat} IS A DAMNED DICTATOR! ARE WE BLIND AND DEAF!?
\stop

\startmarat
THAT'S IT!
\stop

\stage \marat pulls out a pistol! He swings the weapon high above his head -
striking a familiar pose...

\stage The \convention cowers for their lives!

\stage \marat firmly plants the barrel into the side of his head.

\startmarat
I SWEAR TO FUCKING GOD, IF ONE MORE PERSON CALLS ME A DICTATOR, I WILL BLOW MY
FUCKING BRAINS OUT!
\stop

\stage The \convention erupts in fury!

\startmarat
I SWEAR TO GOD I'LL FUCKING DO IT! YOU'LL ALL WATCH ME DO IT!
\stop

\stage \barere lunges at \marat to wrestle the gun from him.

\stage The \convention immediately enters into total combat. The gun goes off
amidst the calamity!

\stage \robespierre stands and produces a Roman salute.

\stage The violence halts.

\stage Pause. Silence.

\stage \robespierre[theincorruptible] ascends the tribune.

\stage Pause. Silence.

\startrobespierre
Silence? What is this, I hear?
\stop

\stage Pause. Silence.

\startrobespierre
Citizens. You have unwittingly been brought far from the true question.

By what mischance has the question that, at one time, united all our votes
become the source of such rage and dissension?

\beat

I have heard doubts cast upon the merits of this trial, on the culpability of
the tyrant, on the integrity of our spirit.

I have listened to decriers accuse legislators of being judges and of
representatives being murderers.

\beat

I hear doubts on the legitimacy of this \convention.

I hear sympathies for royalists who threaten us with their tears.

I hear appeals to philosophy excuse delays and impatience met with
persecutions.

And... I have seen the violence that all of this has led us to.
\stop

\stage Pause. Silence.

\startrobespierre
You are not legislators; you are not judges; this is not a trial; and you are
not philosophers.

You are, and can only be, statesmen; tasked with deciding the measures to be
taken for the one supreme interest of the nation: its safety.

I find it astonishing how far our enthusiasm for this sacred duty has fallen.

Indeed, I recall of a time, not long ago, when each of us were bursting with
excitement for having abolished the monarchy - when we all sang the La
Marseillaise to celebrate the birth of our new republic.

And I recall the extraordinary cheer and applause when the matter of punishing
the great enemy of this republic was brought to motion.

\beat

Yet, now we suddenly cast doubts?

Now, we are suddenly interested in respecting the king's inviolability?

Now, we suddenly wish to honor a constitution penned by men who spoke of
liberty as a concession, rather than as a treasure?

\beat

Allow me, citizens, to call you back to the true question at hand: what measure
must we take to unite our newborn republic?

What message must we carve onto the hearts of the French to rally their
contempt for tyrants?

What gesture shall we submit into history that will strike dumb any who remain
partisan to royalty?

\beat

Shall our message be that the punishment of tyrants is a matter worthy of
lengthy deliberation?

That the truth of \louis' crimes remains uncertain?

That the latent sentiments and guilty hopes of every royalist in France is
due proper respect and consideration?

Or, worst of all, that the power entrusted in our nation's representatives
suffers the fatal flaw of being flimsy and insecure?
\stop

\stage Pause. Silence.

\startrobespierre
Citizens, it is important here that you consider the past, for this is, in
fact, the second time that representatives of the nation have opined on the
punishment of \louis[thelast].

Eighteen months ago, when \louis was brought back to us as a fugitive from his
flight to Varennes, I was present to watch my peers in the Constituent Assembly
rally in condemnation for his actions.

Together they spoke in one voice that he should face the most severe
consequences for his treachery.

But shortly thereafter... other ideas began to emerge.

\beat

Sophisms and intrigues. Questions and sympathies. Doubts and scrutiny.

Quietly, these ideas eventually overcame justice and liberty.

And soon, it became criminal to demand that the full severity of the law be
used against the king.

\louis alone was sacred; and any who defied this were ostracised.

Those men who remained loyal to the people were mercilessly accused of
factionalism; of being seekers of anarchy; and, worse yet, of being
republicans.

\beat

I saw the snares being laid by these enemies of liberty.

And thus, when the fatal petition destined for the Champ-de-Mars was brought to
the \jacobins, it was I alone who objected to it.

Despite the virtuous optimism and zeal for liberty abound, I sought to warn of
the dangers that I had observed in the Assembly.

But, alas, I was not heeded, and the blood of two hundred of the best
citizens~- the blood of women and children~- flowed on the altar of France to
protect the sanctity of a king who hates them.
\stop

\stage Pause. Silence.

\startrobespierre
And now, once again, I listen to these very same sophisms build a maze of
useless and ridiculous chicanery; and have watched, once again, as my peers
become clouded by doubts and persecutions.

Yet, the opportunity remains for us to profit from the experience of the past.

Let us, therefore, impose silence on all our prejudices, on all suggestions and
concerns, and let us return with cold reason to our true question: the question
of our message to the people.

\beat

But first, you ask: which people?

The people of Paris, or the people of France?

This is a false dilemma, for I may ask you the same: who, indeed, do you
picture being represented by the referendum you propose?

Do you picture the farmer abandoning his plow to page through the penal code
and deliberate in assemblies?

Do you picture the artisan sacrificing their daily bread to listen to
confounding pleas for clemency?

Is it, perhaps, in your interest to cloud the vigor of these most pure and
virtuous citizens the same way as you have done here?

Is it your intention to use your guise of humanity and philosophy to spread
hesitancy like a disease across the nation?

To manipulate those citizens who are most vulnerable to the crimes of egoism?
\stop

\stage Pause. Silence.

\startrobespierre
Citizens, allow me to remind you the terrible cost that France has paid for the
ego of careless men.

You would recall that it was I alone who fought for many months to defy the
rally of war on Austria, for it was I alone who saw the danger it presented to
France.

I predicted that the only outcome of war would be an opportunity for scheming
intriguers to either restore the throne to its ancient status, or to have it
seized for themselves.

Yet, as before, I was not heeded, and now, foreign invaders march toward Paris
as we speak.

\beat

And, I note, how curious it is that these men who brought this war upon us...

These men who championed war with the most happy assurances, and who have
sought to mask the horror of its consequences...

And these self-same men, who - not long ago - refused to depose the king in
favor of a mere suspension...

That these are the very men who now plea for delay in the king's punishment.

Who now seek to obstruct justice with tedious and unnecessary procedures...

Who now seem desperate to call for a public referendum.

\beat

Why do they ask for this?

What interests could such a measure advance?

What opportunities would confusion and division provide to scheming men in a
nation already fearful for its safety?

\beat

Citizens. I have observed what is going on around us; I have observed the true
causes of the dissension among us; I see clearly the perils of the system that
shall be the doom of our nation.

And I come to you today with another premonition - that which I fear may be my
last... but one that I must record for the preservation of history should the
worst ill-fated omens come to pass.
\stop

\stage Pause. Silence.

\startrobespierre
I predict... that should the question of \louis be brought to a public
referendum... then we shall see France descend into a civil war.
\stop

\stage The \convention recoils in horror!

\startrobespierre
The referendum that has been proposed is designed to subvert the authority of
the \convention.

Its delay shall provide dissenters the opportunity to requestion everything -
to instill doubts and prejudices amongst the people, and to provoke a civil war
so that its fury may compound with the scourge of foreign invaders.

All so that these ambitious intriguers may lay claim to the ruins of France and
the bloody corpses of its defenders.
\stop

\startbrissot
WHAT!? OUTRAGEOUS! OUT(RAGEOUS) --!?
\stop

\startleftmajority
-- SILENCE! SI-I-ILENCE!
\stop

\startbrissot
AM I NOT A MAN AND A BROTHER!?
\stop

\stage Pause. Silence.

\stage \brissot can feel death.

\startrobespierre
If you would forgive me, I notice a striking resemblance with a tale passed to
us from the ancient Greeks.

\beat

The tragic tale of poor, king Sisyphus.

\beat

The tale of a king who wished to defy death, and, for this transgression, was
condemned by the gods to forever roll a large boulder up a mountain.

And each day, Sisyphus would eventually break from exhaustion, causing the
boulder to roll down the mountain - returning to the plain below.

And yet, each morning, he returns to his duty.

Each day. Breaking. From. Exhaustion.
\stop

\stage Pause. Silence.

\startrobespierre
It has been said that many in Paris feel as if they must suffer this fate.

Indeed, I too share sympathy for poor, king Sisyphus.

But I feel like no king, and I feel this way for a reason, do I not?

\beat

No. Alone, I am not a king.

I am not a judge; I am not a legislator; I am not a philosopher; and I am not
Sisyphus.

\beat{with hate}

I am a montagnard.

A wretched man - struck blind and deaf - with no knowledge but earth and
gravity~- cursed to forever scale the ground beneath my feet in search of
virtue.

And each day, with my eyes gouged and ears punctured, I break myself on the
mountain...

I do so because my labor is my truth.

My labor is both the {\emph will} of the people and the {\emph will} of the
supreme being.
\stop

\stage Pause. \robespierre takes a deep breath.

\startrobespierre
I shall remind you, citizens, that we... we are to be better than kings.

\beat

We are to be a {\emph nation}!

\beat

Alone, I am the montagnard; but together...!
\stop

\starttheincorruptible
WE. SHALL BE. THE MOUNTAIN!
\stop

\stage \themountain rises with incredible ferocity - chants of praise shower
him.

\stage \robespierre pictures himself as Cicero. A childhood dream has come
true.

\stopcomponent
