\startcomponent the_mountain_aux

\stage \robespierre stands and produces a Roman salute.

\stage The shouting halts.

\stage All watch as \robespierre ascends to the tribune.

\stage Pause. Silence.

\startrobespierre
Silence? What is this I hear?
\stop

\stage Pause. Silence.

\startrobespierre
In the stories of the ancient Greeks, we are told the tale of poor, king
Sisyphus: the regent who wished to defy death, and, for his transgression, was
condemned by the gods to forever roll a large boulder up a mountain.

And each day, Sisyphus would eventually break from exhaustion, causing the
boulder to roll down to the plain below.

And yet, each morning: he must return to his duty.

Each day. Breaking. From. Exhaustion.
\stop

\stage Pause. Silence.

\startrobespierre
It has been said that many in Paris feel as if they must suffer this fate.

Indeed, I too sympathize with poor, king Sisyphus.

But I feel like no king, and I feel this way for a reason, do I not?

\beat

No. Alone, I am not Sisyphus, but a petty montagnard.

A wretched man with no understanding but earth and gravity~- destined to
forever scale the ground beneath my feet in search of virtue.

\beat

Each day. With my eyes gouged and ears punctured. I break myself on this
mountain.

I do so because my labor is my truth.

My labor is both the will of the people and the will of the supreme being.
\stop

\stage Pause. Silence.

\startrobespierre
I shall remind you, citizens, that we... We are to be better than kings.

\beat

We are to be a {\emph nation}!
\stop

% TODO: his speeches were generally much longer.

\starttheincorruptible
Alone. I am the montagnard. But together. WE. SHALL BE. THE MOUNTAIN.
\stop

\stage \themountain rises in riotous solidarity.

\stage \robespierre, seeing his newly won majority, pictures himself as a Roman
consul~- a childhood dream has come true.

\stopcomponent
