\startcomponent three_body_problem

\scene

\open \eliza is alone at a desk. A small fire keeps the room warm despite the
wind howling beyond the walls.

\stage \eliza is organizing the documents she has collected, arranging them in
the order she plans to read them.

\stage \eliza chooses a document and begins.

\direction \condorcet is revealed seated in a rocking chair. He writes by the
light of the fire.

\stage Enter \sophie.

\startsophie
\condorcet[c]...
\stop

\stage \condorcet jumps to his feet! Then collapses due to a bad knee.

\stage \sophie rushes to him!

\startcondorcet
{\WORD\sophie}!
\stop

\stage \sophie spots what must be \condorcet's cane and helps him stand.

\stage \condorcet embraces her and cries.

\startcondorcet
\sophie...! Oh, \sophie...

\beat

You shouldn't have come!

How dare you force me to fear for your safety!
\stop

\stage \condorcet begins to sob as \sophie is... strangely paralyzed~- as if
experiencing stage freight to an audience of one.

\startsophie
I wish you would... would understand me when I tell you that the world outside
these walls is so much less severe than you have led yourself to believe.
\stop

\startcondorcet
WHAT!? How could you possibly think that!?

Are... have you have no sympathy for \brissot, or, or... \olympe, or...?
\stop

\stage \sophie is furious~- yet, remains silent.

\stage \condorcet drops to the floor.

\startcondorcet
THEY KILLED {\WORDS \brissot}! THOSE MONSTERS KILLED {\WORDS \brissot}!

\beat{with hate}

Those {\emph fucking} fiends!

\beat{begins pacing}

That {\emph stupid} \marat...!

That cowardly, envious, rat who conducts the savage cries of the crowd so that
he can have the pleasure of watching the goring of citizens from a distance!

How could nature {\emph produce} such a mind so insane as his?

\beat

And that corrupt \danton...!

How... how could a man who is so open to reason decide to use it to sell
decrees!?

Selling {\emph decrees}, \sophie! Selling DECREES!

\beat

You {\emph cannot} sell decrees! It's against the very purpose of the law!

The laws are not bought and sold!

They are {\emph constructed} from our {\emph reason}!

From simple, irrefutable axioms!

OUR NATURAL RIGHTS!

\beat

To think that such a man would {\emph willingly} butcher the public reason for
something as petty as personal wealth!

Fishing in the pond of his own estate while he listens to the world around him
starve!

What the {\emph fuck} does he hope to catch!?
\stop

\stage \condorcet is breathing heavily.

\startcondorcet
\beat{with a hiss}

And {\emph lastly}, that...
\stop

\stage \condorcet freezes... \sophie is terrified...

\stage \condorcet gasps! \sophie jumps!

\startcondorcet
I think I understand...
\stop

\stage \condorcet gasps, again! \sophie jumps!

\startcondorcet
\robespierre...

\beat

\robespierre, he... discovered that French... is indistinguishable from {\emph
		theater}.

\beat

And... and he... abused~- no, {\emph weaponized} this fact...!

\beat

But... w-why? What could compel...?
\stop

\stage Pause.

\startcondorcet
No... No, no... No, no, this... No.
\stop

\stage \sophie's alarms start going off.

\startcondorcet
\beat{like a mad sheep}
NO! NO-O-O! NO-O-O! NO-O-O-O!

\beat{catches his breath}

Don't you see!? DON'T YOU SEE!?

The revolution has been turned into... into... into a {\emph religion}!

A RELIGION, {\WORDS\sophie}...!

And \robespierre installed himself as its {\emph priest}!

\beat

That's what he is, \sophie! A priest! He's a PRIEST!

\beat{like a mad sheep}
HE'S A PRIEST!

HE'S A PRIEST!

HE'S A PRIEST!

HE'S A PRIEST!

HE'S A PRIEST!

HE'S NOTHING BUT A PRI-E-EST!
\stop

\stage \condorcet begins screaming!

\startcondorcet
WHY DOESN'T ANYONE UNDERSTAND ME-E-E~--!?
\stop

\startsophie
-- E-E-ENOUGH!
\stop

\stage Pause. The whispering of winds.

\startsophie
\beat{tears, visibly shaking}
Such tragedy that the mind cannot observe itself!

For yours cannot see how far it has changed!
\stop

\startcondorcet
What? No, that's not~--
\stop

\startsophie
-- Such {\emph tragedy}! That the mind cannot observe others!

For yours cannot see that you are no longer the man I love!
\stop

\startcondorcet
WHAT!? NO! NO, NO! That can't be~--!
\stop

\startsophie
-- SUCH {\emph TRAGEDY}! That the mind is not the master of its fate!

For yours cannot control the laws that govern nature!
\stop

\stage \condorcet painfully returns himself to his feet.

\startsophie
My dear, \condorcet[c]... You are too awfully aware of a recurring truth: that
all enlightenments... are merely partial.

\beat

And I fear that I must tell you that your own enlightenment has reached its
limit.

Thus, I feel it as my duty to inform you that your mind is due... for another.

\beat

O, woe! That you are so convinced that just because you spend of all your time
reflecting on your thoughts that you must therefore know its true function!

O, woe! That you are so prejudiced by this reflection that you have become
blinded to an obvious truth!

\beat

At present, you know your mind as consisting of two estates in a perfect,
harmonious dialogue.

Picturing your thoughts like the forces that compel the moon to orbit the
earth...
\stop

\startsplit

\startcondorcet
Or the earth to orbit the sun!
\stop

\split

\startsophie
And the earth to orbit the sun!
\stop

\stopsplit

\startsophie
The first estate! The estate of \voltaire! \condorcet[therevolutionary]!

The great composer of your worldlines!
\stop

\stage \condorcet gasps! \sophie... is eloquent~- she has rehearsed this many,
many times.

\startsophie
The second estate! The estate of \leibniz! \condorcet[thephilosopher]!

The ingenious constructor of your worldcircles!
\stop

\stage \condorcet gasps! Truly... enchanted.

\startsophie
But... the truth that you have failed to see...

Is that there is yet... a {\emph third}!
\stop

\startcondorcet
Who does the third estate belong to?
\stop

\startsophie
My love... my love, you know so well who it belongs to.

\beat

The third estate... belongs to everyone else.
\stop

\stage \condorcet jumps!

\startsophie
How is this possible, you ask?

Well. The answer is simple! You may fail to believe it could be so simple given
how much anguish this problem has caused you!

\beat

You complain that French cannot be distinguished from theater?

Why, this is true of all languages!
\stop

\stage \condorcet gestures to interrupt~--

\startsophie
-- Oh, except mathematics, you retort?

\beat

Yet, is it not by reason that our emotions are felt?

No matter how unspeakable, or alien... or passionate that reason may be
manifest?

Is it not still by reason alone that it is felt?

\beat

Recall, my dear \condorcet[c], that our gestures are given meaning by their
grammars~- and this meaning-making is performed by the self-same function that
allows our alphabets to emerge from the sounds of our voices.

\beat

And recall! The axiom that reason is independent of the speaker!

Thus, reason {\emph is} that which is shared between all forms of language.

\beat{embraces \condorcet}

Including the ones formed through mutual sympathy.
\stop

\stage \condorcet is finally comfortable.

\startsophie
But I know you, \condorcet[c]. I know your mind demands, not theater, but
proof.

\beat

And... I know where proof has taken you.

I've seen what it has done to your courage.

I've listened to what it has done to your soul.

I see, I hear... and I {\emph feel} the terror that haunts you in the night.

And I... know this terror's {\emph true} name.
\stop

\stage Pause. \condorcet knows precisely what she's referring to.

\startsophie
You call it... strategy.
\stop

\stage Pause. The whispering of winds.

\startsophie
And yes, my dear \condorcet[c], we have arrived at our final destination~- the
conclusion of my evidence.

\beat

Observe that the curse of strategy is, by its definition, the self-same
function that allows our minds to reason about our observations of others...

Therefore, the curse of {\emph infinite} strategy implies the existence of an
everlasting source of empathy for those we observe!
\stop

\stage Pause.

\stage \condorcet gasps! He collapses to the floor!

\startcondorcet
No... There exists... an {\emph equivalency}!
\stop

\stage Pause.

\stage \condorcet begins to break into tears...

\stage Then sobs...

\stage Then suddenly...

\startcondorcet
Eureka...! EUREKA-A-A! EUR-R-R-E-E-E-K-A-A-A!
\stop

\stage \sophie stands in triumph.

\startsophie
Where is he who, faithful to reason and nature, finds peace in the comforts of
domestic virtues?

Where is he who always reserves a part of his soul for the self-reflection that
is the source of all his wonders?

Where is HE? Who's education has taught his heart how to peer so deeply into
the astonishing marvels of the human mind?

\beat{deep breath}

WHERE IS HE? Who's infinite love for humanity is the boundless canvas for which
all his imagination, reason, and hope may be inscribed into the eternal chain
of human destiny?
\stop

\stage \condorcet stands up and looks deep into \sophie's eyes.

\startsophie
Now. Prove it to me.
\stop

\stage \sophie embraces \condorcet~- doing her best to comfort him.

\direction Lights down on \sophie and \condorcet.

\stage \eliza looks up from the documents~- puzzling on something...

\stage Pause.

\stage \eliza takes a deep breath~- gone.

\stopcomponent
