\startcomponent trial_of_louis_xvi

\scene

\open Winter 1792. The Salle du Man\`ege. The \convention[national] has
gathered to begin the trial of \louis[xvi]. In attendance are \condorcet,
\brissot, \paine, \guadet, \gensonne, \vergniaud, \robespierre, \marat,
\danton, and the president of the assembly, \barere.

\stage While the \convention engages in idle chat, a \guard arrives to speak
into the ear of \barere.

\startbarere
Citizens, quiet.
\stop

\stage Pause.

\startbarere
I have been informed that \louis has arrived. He stands at the gates of the
Salle.
\stop

\stage Pause.

\startbarere
Representatives...

In this moment, you exercise the right of national justice.

You have made yourselves answerable to the citizens of France, whom expect from
you the wise and firm conduct which is to be pursued on this momentous
occasion.

\beat

Europe observes you. History records your thoughts and your actions.

Incorruptible posterity will judge you with inflexible severity.

Let our actions here, in this sitting, bestow our legacy with a great lesson to
kings, and allow it to set a useful example for the emancipation of nations.
\stop

\stage Enter \louis, escorted by two \guards.

\stage \louis stands behind a bar that has been constructed for the trial.

\stage Exit the \guards.

\startbarere
\louis. The French Nation accuses you of crimes against its welfare and
security.

The National Convention has decreed on the third of December, that you shall be
judged by them.

We are prepared to now read to you the crimes laid to your charge.

\beat

You may sit down.
\stop

\stage \louis does not defy him. He sits.

\startbarere
\louis. The French Nation accuses you of having committed a multitude of crimes
against her freedoms in the establishment of your tyranny.

\beat

You have, on the twentieth of June, seventeen eighty-nine, at the time of the
Estates-General, attempted to suspend the sovereignty of the people by
threatening violence against the assembly of the third estate.

This is evidenced in the oath established in the tennis-court of Versailles by
the members of the Constituent Assembly.

On the twenty-third of June, you surrounded these representatives with troops
for the purpose of dictating two royal decrees designed to subvert the liberty
of the nation through the dissolution of this assembly.

\beat
What have you to answer?
\stop

\startlouis
No laws existed at the time to prevent me from acting as the sovereign of the
French.
\stop

\stage The \convention grumbles.

\startbarere
You ordered the army to march against the citizens of Paris.

Your satellites spilt the blood of many such citizens.

Despite this, you refused to retreat this army until the taking of the
Bastille, with which it was announced the popular insurrection against your
oppression.

\beat

What have you to answer?
\stop

\startlouis
At that time, I was the master of the troops and was within my rights to order
their march.

But I had no intention of spilling blood.
\stop

\stage The \convention protests. With a glare, \barere silences them.

\startbarere
You had taken an oath, at the Federation of the fourteenth of July, seventeen
ninety, that you have failed to keep.

With wealth in the millions, you attempted, on numerous occasions, to corrupt
officials with bribes in the design of poisoning public opinion and to excite
the provinces into counter-revolutionary movements.

\beat

What have you to answer?
\stop

\startlouis
I do not recollect any such attempts of corruption. As for my oath, the
accusations are contradicted by my acceptance of the constitution the following
year.
\stop

\startbarere
On the twenty-first of July, seventeen ninety-one, you took flight from the
Tuileries using a false passport.

You left behind a letter, wholly in your own hand-writing, to protest the
constitutional articles you swore your oath to.

In the time following this, you ordered that none of your ministers were to,
under any circumstance, sign any act passed to them by the Legislative
Assembly, and you forbade the minister of justice from delivering the seals of
state.

\beat

What have you to answer?
\stop

\startlouis
On the latter, I have no knowledge whatsoever. As for my visit to Varennes, I
stand by my address to the Constituent Assembly that I had been misled by my
ministers.
\stop

\startbarere
Upon your detention at Varennes your executive powers had been summarily
suspended.

Yet, in spite of this, you engaged in a conspiracy to suppress the freedoms of
the people through a show of violence on the seventeenth of July at the Champ
de Mars.

A letter, written in your hand in seventeen ninety, and addressed to
\lafayette, proves that you had entered into a criminal coalition that ties you
to this conspiracy.

\beat

What have you to answer?
\stop

\startlouis
What happened on the seventeenth of July had nothing to do with me.

I was not involved in any way and had no knowledge of it until the following
morning.

The insinuation that my letters to \lafayette\ - written a year prior to the
tragedy - serve as proof of my involvement is absurd.
\stop

\stage The \convention protests. \barere's glare fails this time.

\startbarere
Silence!
\stop

\stage The \convention simmers into grumbles.

\startbarere
Your brothers, including Leopold of Austria, who are enemies of the state, made
arrangements with the emigrants to rally under their banners.

They raised regiments, took up loans, and ratified alliances in your name.

This is evidenced in a letter addressed to you, signed by your two brothers,
for which it reads:

\beat{holds up letter}

"We are two here who now make but one: one in sentiment, one in principles, one
in zeal to serve you."
\stop

\stage The \convention erupts with disgust. \barere continues in spite of it.

\stage \louis maintains his dignified composure.

\startbarere
"We keep our silence for fear of compromising you, but we shall speak as soon
as we shall be certain of general support, and that moment draws near.

If we are spoken to on the part of those people, we will hear nothing.

But in your part, we shall listen dutifully.

We await to hear your correspondence. Remain vigilant for your safety.

We only exist to serve you."
\stop

\startbarere
What have you to answer?
\stop

\stage \louis takes a deep breath.

\startlouis
I disowned all the actions and proceedings of my brothers - exactly as
prescribed by the constitution.

This note you read... I know nothing of it.
\stop

\stage The \convention continues their shouting.

\startbarere
The interior of...

THE INTERIOR OF...
\stop

\stage The \convention settles down.

\startbarere
The interior of the state had become corrupted by fanatics, and you swore
yourself to them as their protector.

In doing so, you manifested your intention of restoring the monarchy to its
ancient status - what have you to answer?
\stop

\startlouis
I cannot answer to this! I know nothing of such a project!
\stop

\startbarere
The Legislative Assembly had passed a decree on the twenty-ninth of January
against the fanatical priests, yet you had suspended its execution - what have
you to answer?
\stop

\startsplit

\startbrissot
Here, here!
\stop

\blank[line]

\startguadet
Here, here!
\stop

\split
\leavevmode

\startgensonne
Here, here!
\stop

\blank[line]

\startvergniaud
Here, here!
\stop

\stopsplit

\startlouis
The constitution reserved to me the right of veto.
\stop

\startbarere
The troubles in France had escalated; the minister of justice announced that he
had no legal means to arraign the guilty.

The Legislative Assembly enacted a decree to remedy this, yet you likewise
suspended its execution - what have you to answer?
\stop

\startlouis
The constitution reserved to me the right of veto.
\stop

\startbarere
You employed Swiss guards at the Tuileries despite the constitution forbidding
this.

The Legislative Assembly ordained you to dismiss them, yet you refused to do so
- what have you to answer?
\stop

\startlouis
The employ of private mercenaries does not contradict the articles of the
constitution.
\stop

\startsplit

\startlouis
As for the latter, the constitution reserved to me the right of veto.
\stop

\split
\leavevmode

\startmarat
Liar! LIA-A-AR!
\stop

\stopsplit

\startrightminority
Liar! Liar! Liar!
\stop

\startbarere
You had personally reviewed the station of the Swiss guards at the Tuileries on
the tenth of last August at five o'clock in the morning.

It was these same guards who fired first upon the citizens who arrived in
protest that day.
\stop

\startlouis
I reviewed all of my troops that morning.

The constituted authorities were with me, the department, the mayor, the
municipality - all were with me!

I even invited them to join my family for breakfast!
\stop

\startbarere
Why did you draw troops to the Tuileries?
\stop

\startlouis
All the constituted authorities saw that the castle was in danger - that my
family was under threat!

I had the right to defend myself!
\stop

\startbarere
You have caused the blood of the French to shed.
\stop

\stage The \convention begin to jeer.

\startlouis
No! That's a lie! A LIE!
\stop

\startmarat
TO THE SCAFFO-O-OLD! TO THE SCAFFO-O-OLD! DEATH FOR THE TYRANT!
\stop

\startlouis
NO! No! I - I... I demand my counsel! I DEMAND COUNSEL!
\stop

\startbarere
\beat{to the \convention}

The questioning is done with.

\beat{to \louis}

\louis. Is there anything you wish to add?
\stop

\stage The \convention begins deliberation.

\startlouis
I... would like to request counsel...
\stop

\startbarere
I invite you to withdraw to the conference hall. The assembly must deliberate.
\stop

\stage Enter two \guards.

\startmarat
GUILTY! HE IS GUILTY! GUARDS! DRIVE YOUR BAYONETS INTO HIS NECK!
\stop

\startbarere
There has been no motion for judgement!
\stop

\stage Exit \louis and the \guards.

\startthuriot
HE SHOULD ASCEND THE SCAFFOLD!
\stop

\startsplit

\startguadet
HE DESERVES COUNSEL!
\stop

\split
\leavevmode

\startduhem
TO HELL WITH COUNSEL!
\stop

\stopsplit

\startbrissot
YOU MONSTERS DEBASE YOURSELVES!
\stop

\stage The \convention enters an uproar.

\stage \barere rings a bell.

\startbarere
SILENCE, YOU HEATHENS!
\stop

\startpetion
\barere[citizen], I request the floor!
\stop

\startbarere
Granted.
\stop

\startpetion
\beat{from his seat}

What the hell is the matter with this assembly?

What question is there in a defendant's request for counsel?

Are not most of us trained in law? Whatever due process can there be without a
defense?

How is it possible for such a simple question to create such division amongst
ourselves?

\beat

So, the late king demands counsel, so what? I say no one can refuse him unless
it is their will to attack all the principles of humanity at once.

I put to motion the question - the simple question: "should \louis receive
counsel?"
\stop

\stage The \leftminority applaud him.

\startthuriot
I say he ascend the scaffold! Put that to motion!
\stop

\stage The \rightminority applaud him.

\startduhem
I request the floor, \barere[citizen]!
\stop

\stage \barere gestures in the affirmative.

\startduhem
I demand the immediate sentence of \louis by nominal call.
\stop

\stage The \rightminority applaud him.

\startduhem
I put to motion the question - the simple question: "shall he suffer death, or
shall he not?"

I say, let him be executed, and after that publish his defense.
\stop

\stage The \rightminority give him a standing ovation.

\startsplit

\startlajuinais
I OPPOSE this barbarous --
\stop

\split
\leavevmode

\startmarat
-- He speaks without the floor! TO THE ABBEY WITH HIM! LOCK HIM UP!
\stop

\stopsplit

\stage \barere rolls his eyes.

\startbrissot
Hypocrite! \marat is a shameless despot!
\stop

\stage The \convention erupts!

\startbarere
SILENCE! SI-I-ILENCE!
\stop

\stage \barere gestures to hear \lajuinais.

\startlajuinais
\beat{groans}

I oppose this barbarity. The time for bloodthirsty men is ancient history.

Let the fate of \louis be that which public safety requires.

To this end, I see it fit to deny any opportunity for the conspirators
responsible for the tenth of August to judge this man.

These men clearly have no patience for justice.

I say that the question of judgement should therefore be decided by the public.
\stop

\stage An uproar from the \rightminority! The \leftminority give him a standing
ovation.

\stage A member of the \rightminority gets up from his seat and proceeds to
attack \lajuinais.

\stage Members of the \leftminority dive into the calamity to restrain the
attacker.

\stage Warfare ensues.

\direction Lights down on all but \condorcet.

\stage Enter \voltaire.

\startvoltaire
Professor. Do we live in the best of all possible worlds?
\stop

\stage Enter \leibniz.

\startleibniz
I suspect you ask in bad faith.
\stop

\startvoltaire
What would give you that impression?
\stop

\stage Pause.

\startvoltaire
What say you, \leibniz? Yay, or nay? Cast your vote.
\stop

\startleibniz
It is a matter of reason, \voltaire. There is no nay.
\stop

\startvoltaire
Is reason not the purpose of our votes? Of these debates?
\stop

\startleibniz
But you ask about the nature of the world and this encompasses all things that
exist.

It is, therefore, not contingent on any particular worldline.
\stop

\startvoltaire
I disagree. I think it has everything to do with my worldline.
\stop

\startleibniz
This is a common misunderstanding.

The goodness of our world over any possible alternative is not a feature of the
world in part - it is a feature of the world as a whole, across all of time.

The decision made by God to select for a world that permits particular evils,
such as the violence we see today, is beyond the comprehension of mere
creatures.
\stop

\startvoltaire
What decision function do you suppose God uses?
\stop

\startleibniz
Your question is not valid. Such a function is beyond human understanding.
\stop

\startvoltaire
Is it?

Clearly it must be a trivial one since nothing we can do about it seems to
matter.
\stop

\startbarere
\condorcet.
\stop

\startleibniz
A decision function, as we have defined the term, only applies to social
decisions - not to the decisions of --
\stop

\startbarere
-- \condorcet!
\stop

\direction Lights up. The hall has emptied aside from \barere and \condorcet.

\startbarere
Everyone has left for recess. Are you really going to sit there in silence for
all of two hours?
\stop

\startcondorcet
Oh. I... I suppose not. I apologize.
\stop

\startbarere
\beat{chuckles}
Don't apologize to me. Your silence is no burden.
\stop

\stage Exit \condorcet.

\stage \barere rifles through the documents on his desk.

\stage Pause.

\stage \barere breaks into sobs, then suffers to silence himself.

\stopcomponent
